---
title: "R Notebook"
output: html_notebook
---

Import libraries
```{r}
suppressPackageStartupMessages(library(metaseqR))
suppressPackageStartupMessages(library(DESeq2))
```


#Figure 2
A: PCA
B: "Rank sum": cpm of top 20 species in unblocked, same species in blocked
C: DE- DESeq2, excluding heme miRs
D: log(CPM) of raw counts


Functions
```{r}
#Create metadata and DESeq2 object, then run DESeq2

RunDESeq2<- function(counts,meta, collapseRep=TRUE) {
  #create DESeq2 object- just based on samplegroup (blockerstatus)
  #will add samples to model later (i.e., pairwise analysis)
  dds<- DESeqDataSetFromMatrix(countData=counts, 
                               colData=meta, 
                               design=~sampleGroup)
  #Collapse technical replicates using internal DESeq2 function
  if (collapseRep ==TRUE){
    dds<- collapseReplicates(dds, dds$sample)
  }
  
  #Create model that takes into account sample differences
  ddsMF<- dds
  design(ddsMF)<- formula(~Patient + sampleGroup)
  
  #run DESeq2
  dds<- DESeq(ddsMF)
  
  return(dds)
}

#Get DE results
GetDEResults<- function(dds){
  res<- results(dds)
  res<- na.omit(res)
  resOrdered<- res[order(res$padj),]
  print(nrow(resOrdered[resOrdered$padj<0.05,]))
  print(nrow(resOrdered[resOrdered$padj<0.01,]))
  return(resOrdered)
}

MyPCA<- function(dds, metadata, PC1Variance, PC2Variance, Groupcolors,miny, maxy, minx, maxx) {
  #Normalized (variance stabilized transformation), then run PCA
  vsd <- varianceStabilizingTransformation(dds, blind=TRUE)
  PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)
  PCA$Patient<- metadata$Patient
  
  #plot
  p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group, shape=Patient))+
  geom_point(size=8)+
  scale_color_manual(values=Groupcolors)+
  scale_shape_manual(values=c(15,20,43,18))+
  theme_classic()+
  ylim(miny,maxy)+
  xlim(minx,maxx)+
  ylab(paste("PC2: ", PC2Variance, "% Variance", sep =""))+
  xlab(paste("PC1: ", PC1Variance, "% Variance", sep =""))+
  #geom_text(aes(label=Sample),hjust=.5, size= 5,vjust=-1.2, color="black")+
  theme(axis.text.y   = element_text(size=12, colour = "black", face="bold"),
        axis.title.y  = element_text(size=14, colour = "black", face-"bold"),
        axis.title.x  = element_text(size=14, colour = "black", face="bold"),
        axis.text.x  = element_text(size=12, colour = "black", face-"bold"),
        legend.background = element_rect(fill=NA, size=1, 
                                         linetype="solid", color="black"),
        legend.title=element_blank(),
        legend.text = element_text(size=14, face="bold"),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"),
        legend.position = c(.9,.85),
        panel.border = element_rect(colour = "black", fill=NA, size=2))
  return(p)
}

MyPCA_TechNotCollapsed<- function(counts, meta,PC1, PC2, miny, maxy, minx, maxx){
  #Rename samples
  TechNotCollapsed<- counts
  colnames(TechNotCollapsed)<- c(paste("A.0.", 1:3,sep=""), paste("B.0.", 1:3,sep=""), "C.0", "D.0", paste("A.1.", 1:3,sep=""), paste("B.1.", 1:3,sep=""), "C.1", "D.1")


  #create DESeq2 object
  dds_TNC<- DESeqDataSetFromMatrix(countData=TechNotCollapsed, 
                             colData=meta, 
                             design=~sampleGroup)


  #Create model that takes into account sample differences
  ddsMF_TNC<- dds_TNC
  design(ddsMF_TNC)<- formula(~Patient + sampleGroup)

  #run DESeq2
  dds_TNC<- DESeq(ddsMF_TNC)
  
  #Run through MyPCA function to plot
  p<- MyPCA(dds_TNC, meta, PC1, PC2, Groupcolors, miny, maxy, minx, maxx)
  return(p)
}
```

#Read in data and downsample

```{r}
#readwd<-"Z:/Channing_miRNA_Blocker_Test/BlockerProject/BlockerProject/RawData/"
#countsInput<- read.csv(paste(readwd, "miRNA_CLmiRDeep2Results_AllSamples_DefaultmiRDeep2Settings.csv", sep=""))
countsInput<- read.csv("C:/Users/Jenna/Documents/BlockerProject/miRNA_CLmiRDeep2Results_AllSamples_DefaultmiRDeep2Settings.csv")
rownames(countsInput)<-countsInput$X
countsInput<- countsInput[,-1]
countsInput<- countsInput[,c(1,3,5,7,9,11,13,15,2,4,6,8,10,12,14,16)]
colnames(countsInput)<- c(paste(c("A:low", "A:medium", "A:high", "B:low", "B:medium", "B:high", "C", "D"), "-Unblocked",sep=""),
                          paste( c("A:low", "A:medium", "A:high","B:low", "B:medium", "B:high", "C", "D"), "-Blocked",sep="" ))
```

Downsample to common depth
```{r}
counts<- downsample.counts(countsInput, set.seed(42)) #9,989,167 lowest B.570.0
```

#Figure 2A: PCA

```{r}
#Need to detach DESeq (need to use DESeq2)- used for metaseqR
detach("package:metaseqR")
detach("package:DESeq")
```

General plotting variables
```{r}
Groupcolors<- c("firebrick3","turquoise3")
```

prep data for DESeq2, create metadata and DESeq2 object, then run DESeq2 and PCA. Here, technical replicates are NOT collapsed, since this dds object will be used for plotting PCA
```{r}
#convert counts to integer
counts<-as.data.frame(apply(counts, 2, as.integer))
rownames(counts)<- rownames(countsInput)

#rename rows, removing input info
colnames(counts)<- c(rep("A.0", 3), rep("B.0", 3), "C.0", "D.0", rep("A.1", 3), rep("B.1", 3), "C.1", "D.1")

#Create metadata- will be used in all analyses- when collapsing technical replicates and when leaving separate:
sampleGroup<- as.factor(c(rep("Unblocked", 8), rep("Blocked", 8))) #set blocker status
sample<- colnames(counts) #this column is used to collapse technical replicates
Patient<- as.factor(c(rep("A", 3), 
                      rep("B", 3), 
                      "C", "D", 
                      rep("A", 3), 
                      rep("B", 3), 
                      "C", "D")) #this column is to control for sample to sample differences
meta<- data.frame(sample, sampleGroup, Patient)


#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData = counts,
                             colData=meta,
                             design= ~sampleGroup)



#Normalize (variance stabilized transformation) 
vsd <- varianceStabilizingTransformation(dds, blind=TRUE)

#run PCA
  PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)
  PCA$Patient<- meta$Patient
  
  #plot
p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group, shape=Patient))+
          geom_point(size=8)+
          scale_color_manual(values=Groupcolors)+
          scale_shape_manual(values=c(15,20,43,18))+
          theme_classic()+
          ylim(-20,20)+
          xlim(-22, 35)+
          ylab(paste("PC2: ", 25, "% Variance", sep =""))+
          xlab(paste("PC1: ", 30, "% Variance", sep =""))+
          #geom_text(aes(label=Patient),hjust=.5, size= 5,vjust=-1.2, color="black")+
          theme(axis.text.y   = element_text(size=12, colour = "black", face="bold"),
                axis.title.y  = element_text(size=18, colour = "black", face="bold"),
                axis.title.x  = element_text(size=18, colour = "black", face="bold"),
                axis.text.x  = element_text(size=12, colour = "black", face="bold"),
                legend.background = element_rect(fill=NA, size=1, 
                                                 linetype="solid", color="black"),
                legend.title=element_blank(),
                legend.text = element_text(size=16, face="bold"),
                legend.key.size = unit(.4, "cm"),
                legend.spacing.y = unit(0, "mm"),
                legend.position = c(.88,.82),
                panel.border = element_rect(colour = "black", fill=NA, size=2))
          

```

#Figure 2C: DESeq2

#First: supplemental figure justifying which extra miRs to remove- in same family as targets, also affected by blockers

```{r}
#Entire list of all species in target family
Target_92<- rownames(counts[grep("miR-92", rownames(counts))[5:9],])
Target_451<- rownames(counts[grep("miR-451", rownames(counts))[11],])
Target_486<- rownames(counts[grep("miR-486", rownames(counts)),])
Target_25<- rownames(counts[grep("miR-25", rownames(counts)),])
TargetFamilies<- c(Target_451,Target_486,Target_92, Target_25)

#Get raw counts of these species
MapTargetFamilies<- counts[rownames(counts) %in% TargetFamilies,]


#Get percentage of total for all 3 counts for all 16 samples 
CPMMapToTargetFamilies<- MapTargetFamilies/colSums(counts) * 1000000

#Collapse into two groups- blocked and unblocked
CPMMapToTargetFamilies$Unblocked<- rowMeans(CPMMapToTargetFamilies[,1:8])
CPMMapToTargetFamilies$Blocked<- rowMeans(CPMMapToTargetFamilies[,9:16])
CPMMapToTargetFamily_collapsed<- CPMMapToTargetFamilies[,17:18]

#Only keep if CPM is somewhat high- greater than 50
CPMMapToTargetFamily_collapsed<- CPMMapToTargetFamily_collapsed[CPMMapToTargetFamily_collapsed$value]
CPMMapToTargetFamily_collapsed$species<- rownames(CPMMapToTargetFamilies)
CPMMapToTargetFamily_collapsed<- melt(CPMMapToTargetFamily_collapsed)


p<-ggplot(CPMMapToTargetFamily_collapsed, aes(y=value, x=species,fill=variable))+
  geom_bar(stat="identity", position="dodge", width=.7)+ 
  theme_classic()+
  ylab("Percent of reads mapping to targets species")+
  xlab("")
  scale_fill_manual("", values=c("Blocked"="firebrick3", "Unblocked"="turquoise3"))
  geom_errorbar(aes(ymin=MeanPercentMapping-sd, ymax=MeanPercentMapping+sd), width=.1, size=1, position=position_dodge(.9),color="black")+
  ylim(0,100)+
  theme(axis.text.y   = element_text(size=12, color="black", face= "bold"),
          axis.text.x   = element_text(size=14, color="black", face="bold"),
          axis.title.y  = element_text(size=14, color="black", "bold"),
          panel.border = element_rect(colour = "black", fill=NA, size=2),
          legend.title=element_blank(),
          legend.position = "none",
          legend.background = element_rect(fill=NA, size=1, linetype="solid", color="black"),
          axis.text.x.bottom  =element_text(size=20))

```


```{r}
#Filtering possible:
        #1: Remove heme miRs from counts dataset
        counts_noHeme<- counts[!(rownames(counts) %in% c("hsa-miR-92a-3p", "hsa-miR-451a", "hsa-miR-486-5p")),]
        
        #2: filter based on cpm- only keep species where cpm > 1 in at least two samples. 
        #filter raw counts based on this filtered cpm list. Use raw counts for DESeq2 later
        counts_cpm<- counts/colSums(counts) *1000000
        
        #Determine how many samples pass filter for each species
        PassFilter<- apply(counts_cpm, 1, function(x) {sum(x>1)})
        
        #remove any species that don't have at least 2 samples passing filter
        cpm_PassFilter_2Samples<- counts_cpm[PassFilter>2,]
        
        counts_noheme_cpmpass<- counts_noHeme[rownames(counts_noHeme) %in% rownames(cpm_PassFilter_2Samples),]
        
        #3: remove heme miRs + species in heme miR families
        counts_noHemeFamily<- counts[!(rownames(counts) %in% TargetFamilies),]

dds_counts<- counts_noHeme
#Run DESeq2- this time collapsing technical replicates
dds<- DESeqDataSetFromMatrix(countData=dds_counts, 
                               colData=meta, 
                               design=~sampleGroup)

#Collapse technical replicates using internal DESeq2 function
dds<- collapseReplicates(dds, dds$sample)
  
#Create model that takes into account sample differences
design(dds)<- formula(~Patient + sampleGroup)

#run DESeq2
dds<-DESeq(dds)

#Get DEResults
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
print(nrow(resOrdered[resOrdered$padj<0.05,])) #filter 1: / filter 1+2: 37 / filter 3: 27
print(nrow(resOrdered[resOrdered$padj<0.01,])) #filter 1: / filter 1+2: 25 / filter 3: 19

#normalize results
rld<- rlog(dds)

#create DESeq2 object- just based on samplegroup (blockerstatus)
  #will add samples to model later (i.e., pairwise analysis)
  dds<- DESeqDataSetFromMatrix(countData=counts, 
                               colData=meta, 
                               design=~sampleGroup)
  #Collapse technical replicates using internal DESeq2 function
  if (collapseRep ==TRUE){
    dds<- collapseReplicates(dds, dds$sample)
  }
  
  #Create model that takes into account sample differences
  ddsMF<- dds
  design(ddsMF)<- formula(~Patient + sampleGroup)
  
  #run DESeq2
  dds<- DESeq(ddsMF)
  
  return(dds)
```



```{r}
#Plot top DE species on heatmap

#set plotting parameters
nspecies<- 100
MinSubtract<- -.2
MaxSubtract<- 1

#Create metadata for collapsed 
meta_collapsed<- meta[c(1,4,7,8,9,12,15,16),]

#get top genes of normalized DESeq2 results
topgenes<- rownames(resOrdered[1:nspecies,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)
colnames(mat)<-paste(meta_collapsed$Patient," (",meta_collapsed$sampleGroup, ")",sep="")

#Set breaks for heatmap
mat_breaks<- seq(min(mat-MinSubtract), max(mat-MaxSubtract), length.out=75)

#Create metadata- to be used to color code to show which group the sample belongs to
metadata<- data.frame(Group=meta_collapsed$sampleGroup, 
                row.names=paste(meta_collapsed$Patient," (",meta_collapsed$sampleGroup, ")",sep=""),
                Sample=meta_collapsed$Patient)

  #plot heatmap
  p<-pheatmap(mat, breaks = mat_breaks, 
            color =colorRampPalette( c("red", "black", "green"), space="rgb")(100),
            show_rownames = TRUE, show_colnames=FALSE,
            annotation = metadata,  
            annotation_colors= list(
              Group=c(Blocked="firebrick3", Unblocked="blue3"),
                      Sample=c(A="goldenrod1", 
                               B="violetred3", 
                               C="skyblue3", 
                               D="darkolivegreen4")),
            fontsize_row = 12,fontsize=12,treeheight_row=0,
            border_color= NA)
```

