---
title: "Improvements with blockers: passing thresholds, impact on variation "
output: html_notebook
---

Import libraries
```{r}
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(metaseqR))
```


#Examine potential improvements to sensitivity/variability with the use of blockers

First:

How many species pass a certain count threshold in blocked vs unblocked samples?

Import and reformat data:
```{r}
#Import
wd<- "C:/Users/jjl71/Documents/BlockerProject/BlockerProject/"
CL_countsInput_Downsampled<- read.csv(paste(wd, "RawData/Downsampled_miRNA_CLmiRDeep2Results_AllSamples_DefaultmiRDeep2Settings.csv",sep=""))

#Reformat
rownames(CL_countsInput_Downsampled)<-CL_countsInput_Downsampled$X
CL_countsInput_Downsampled<- CL_countsInput_Downsampled[,-1]

#Reorder 
CL_countsInput_Downsampled<- CL_countsInput_Downsampled[,c(1,3,5,7,9,11,13,15,2,4,6,8,10,12,14,16)]

#Convert to integer (removes decimal places)
CL_countsInput_Downsampled_Integer<- as.data.frame(lapply(CL_countsInput_Downsampled, as.integer))
rownames(CL_countsInput_Downsampled_Integer)<- rownames(CL_countsInput_Downsampled)
CL_Downsampled<- CL_countsInput_Downsampled_Integer
```

```{r}
#Collapse technical replicates (by just taking mean of counts)
unblockedCounts<- as.data.frame(CL_Downsampled[,1:8])
unblockedCounts$ACollapsed<- rowMeans(unblockedCounts[,1:3])
unblockedCounts$BCollapsed<- rowMeans(unblockedCounts[,4:6])
unblockedCounts<-unblockedCounts[,c(9,10,7,8)]

#get mean counts across all 4 samples
unblockedCountsMeans<- as.data.frame((rowMeans(unblockedCounts)))

#number that pass thresholds:
unblockedPassThresholds<- data.frame(threshold=c(1,5,10,50,100,500), countsPassingThreshold=c(sum(unblockedCountsMeans>1),sum(unblockedCountsMeans>5),sum(unblockedCountsMeans>10),sum(unblockedCountsMeans>50),sum(unblockedCountsMeans>100),sum(unblockedCountsMeans>500)))

#blocked: collapse technical replicates (by just taking mean of counts)
blockedCounts<- as.data.frame(CL_Downsampled[,9:16])
blockedCounts$ACollapsed<- rowMeans(blockedCounts[,1:3])
blockedCounts$BCollapsed<- rowMeans(blockedCounts[,4:6])
blockedCounts<-blockedCounts[,c(9,10,7,8)]

#get mean counts across all 4 samples
blockedCountsMeans<- as.data.frame((rowMeans(blockedCounts)))

#number that pass thresholds:
blockedPassThresholds<- data.frame(threshold=c(1,5,10,50,100,500), countsPassingThreshold=c(sum(blockedCountsMeans>1),sum(blockedCountsMeans>5),sum(blockedCountsMeans>10),sum(blockedCountsMeans>50),sum(blockedCountsMeans>100),sum(blockedCountsMeans>500)))

#combine into one, adding in an extra column for blocker status
unblockedPassThresholds$status<- rep("unblocked",6)
blockedPassThresholds$status<- rep("blocked",6)
passThreshold<-rbind(unblockedPassThresholds,blockedPassThresholds)

#plot
p<- ggplot(passThreshold, aes(x=threshold, y=countsPassingThreshold, color=status))+
             geom_line(size=2)+
  geom_point(size=4)+
  scale_color_manual(values=c("black", "red"))+
  ggtitle("Number of miRNAs passing a count threshold")+
  theme_classic()
p
```

#doing the above, but not taking the mean. still collapse technical, but keep a/b/c/d info. Add in a 4th column in the final table with sample type. Then subtract the #passing threshold for a/b/c/d individually from blocker vs no blocker, then graph. 

whoowhweeee

```{r}
#write (awful) function that takes a column of counts and returns the number of species that pass the 6 threshold already written into the function, bc i am stupid
CountsPassingThreshold<- function(counts){
  return(c(sum(counts>1),sum(counts>5),sum(counts>10),sum(counts>50),sum(counts>100),sum(counts>500),sum(counts>1000)))
}

#Get counts passing threshold for unblocked samples, then transform data- 4 columns: ID (A/b/c/d), threshold, #of species that passed that threshold, and blocker status
unblockedPassThreshold<- as.data.frame(apply(unblockedCounts, 2, CountsPassingThreshold))
rownames(unblockedPassThreshold)<- c(1,5,10,50,100,500,1000)
colnames(unblockedPassThreshold)<- c("A", "B", "C", "D")
#add 4th column: mean of all samples
unblockedPassThreshold$Mean<- rowMeans(unblockedPassThreshold)
unblockedPassThreshold<- as.data.frame(t(unblockedPassThreshold))
unblockedPassThreshold$ID<- rownames(unblockedPassThreshold)
#reformat- melt
unblockedPassThreshold<- melt(unblockedPassThreshold)
colnames(unblockedPassThreshold)<- c("ID", "threshold","speciesPassingThreshold")
unblockedPassThreshold$status<- rep("unblocked", 35)
unblockedPassThreshold$threshold<- as.integer(as.character((unblockedPassThreshold$threshold)))

#Get counts passing threshold for blocked samples, then transform data- 4 columns: ID (A/b/c/d), threshold, #of species that passed that threshold, and blocker status
blockedPassThreshold<- as.data.frame(apply(blockedCounts, 2, CountsPassingThreshold))
rownames(blockedPassThreshold)<- c(1,5,10,50,100,500,1000)
colnames(blockedPassThreshold)<- c("A", "B", "C", "D")
#add 4th column: mean of all samples
blockedPassThreshold$Mean<- rowMeans(blockedPassThreshold)
blockedPassThreshold<- as.data.frame(t(blockedPassThreshold))
blockedPassThreshold$ID<- rownames(blockedPassThreshold)
#reformat- melt
blockedPassThreshold<- melt(blockedPassThreshold)
colnames(blockedPassThreshold)<- c("ID", "threshold","speciesPassingThreshold")
blockedPassThreshold$status<- rep("blocked", 35)
blockedPassThreshold$threshold<- as.integer(as.character((blockedPassThreshold$threshold)))

#combine into one
passThreshold<-rbind(unblockedPassThreshold,blockedPassThreshold)

#reformat so that you have 4 columns: ID, threshold, #species passing in noblocker, #species passing in blocker
PT<-data.frame(passThreshold$ID[1:35], passThreshold$threshold[1:35],passThreshold$speciesPassingThreshold[1:35], passThreshold$speciesPassingThreshold[36:70])
colnames(PT)<- c("ID", "threshold", "SpeciesPT_noblocker", "SpeciesPT_blocker")

#now need to subtract each- i.e., sample A at threshold 1 with no blocker vs with blocker. get those values for all, then plot. 
df<- as_tibble(PT)
df<-df%>% group_by(ID) %>% mutate(BlockerMinusNoBlocker=SpeciesPT_blocker-SpeciesPT_noblocker)
df<- as.data.frame(df)

#plot
p<- ggplot(df, aes(x=threshold, y=BlockerMinusNoBlocker, color=ID))+
             geom_line(size=1)+
  geom_point(size=2)+
  scale_color_manual(values=c("turquoise2", "deeppink1", "goldenrod1", "darkviolet","black"))+
  ggtitle("Number of miRNAs passing a count threshold")+
  theme_classic()
p

```


