---
title: "R Notebook"
output: html_notebook
---
Libraries:
```{r}
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(metaseqR))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(dplyr))
windowsFonts("Arial" = windowsFont("Arial"))
```

#Figure 5: improvements with blockers
A: number of species that pass threshold
B-C: improvements in DE

Functions used
```{r}
#Function that takes a column of counts and returns the number of species that pass the given list of threhsolds
CountsPassingThreshold<- function(counts, threshold){
  Passthreshold<-c()
  for (i in 1:length(threshold)){
    Passthreshold[i]<- sum(counts>threshold[i])
    
  }
  return(Passthreshold)
}

#When comparing DE efficiency in blocked vs unblocked, this function used to find species unique to blocked/unblocked and find raw counts for those species. Used for plotting.

GetCounts_UniqueDEmiRNA<- function(BlockedSig, UnblockedSig, SigLevel, RawCounts){
  #Get data frame of sig miRNAs for blocked and unblocked at given significance level
  NBSig<-UnblockedSig[UnblockedSig$padj<SigLevel,]
  BSig<- BlockedSig[BlockedSig$padj<SigLevel,]
  
  #Get DE miRNA specific to blocked and unblocked
  DEInUB<- rownames(as.data.frame(BSig)) %in% rownames(as.data.frame(NBSig))
  DEInB<- rownames(as.data.frame(NBSig)) %in% rownames(as.data.frame(BSig))
  OnlyBlockedDE<- BSig[!DEInUB,]
  OnlyUnBlockedDE<- BSig[!DEInB,]
  
  #Get raw counts for these miRNAs
  counts_onlyblockedDE<- RawCounts[rownames(RawCounts) %in% rownames(OnlyBlockedDE),]
  counts_onlyblockedDE<- counts_onlyblockedDE[,c(1:6,9:14)]
  counts_onlyblockedDE<-counts_onlyblockedDE[order(counts_onlyblockedDE$A.0, decreasing=TRUE),]
  
  counts_onlyunblockedDE<- RawCounts[rownames(RawCounts) %in% rownames(OnlyUnBlockedDE),]
  counts_onlyunblockedDE<- counts_onlyunblockedDE[,c(1:6,9:14)]
  counts_onlyunblockedDE<-counts_onlyunblockedDE[order(counts_onlyunblockedDE$A.1, decreasing=TRUE),]
  
  return(list(counts_onlyblockedDE, counts_onlyunblockedDE))
  
}

GetSigSpecies<- function(SigLevel, res){
        return(res[res$padj < SigLevel,])
}
```

#Read in data and downsample

```{r}
#readwd<-"Z:/Channing_miRNA_Blocker_Test/BlockerProject/BlockerProject/RawData/"
#countsInput<- read.csv(paste(readwd, "miRNA_CLmiRDeep2Results_AllSamples_DefaultmiRDeep2Settings.csv", sep=""))
countsInput<- read.csv("C:/Users/Jenna/Documents/BlockerProject/miRNA_CLmiRDeep2Results_AllSamples_DefaultmiRDeep2Settings.csv")
rownames(countsInput)<-countsInput$X
countsInput<- countsInput[,-1]
countsInput<- countsInput[,c(1,3,5,7,9,11,13,15,2,4,6,8,10,12,14,16)]
colnames(countsInput)<- c(paste(c("A:low", "A:medium", "A:high", "B:low", "B:medium", "B:high", "C", "D"), "-Unblocked",sep=""),
                          paste( c("A:low", "A:medium", "A:high","B:low", "B:medium", "B:high", "C", "D"), "-Blocked",sep="" ))

#Downsample to common depth
counts<- downsample.counts(countsInput, set.seed(42)) #9,989,167 lowest B.570.0

#Convert values to integer
counts<-as.data.frame(apply(counts, 2, as.integer))
rownames(counts)<- rownames(countsInput)
```




#Figure 5A:

Get number of species that pass certain thresholds, plot
```{r}
#Collapse technical replicates (by just taking mean of counts)
unblockedCounts<- as.data.frame(counts[,1:8])
unblockedCounts$ACollapsed<- rowMeans(unblockedCounts[,1:3])
unblockedCounts$BCollapsed<- rowMeans(unblockedCounts[,4:6])
unblockedCounts<-unblockedCounts[,c(9,10,7,8)]

#get mean counts across all 4 samples
unblockedCountsMeans<- as.data.frame((rowMeans(unblockedCounts)))

#number that pass thresholds:
unblockedPassThresholds<- data.frame(threshold=c(1,5,10,50,100,500), countsPassingThreshold=c(sum(unblockedCountsMeans>1),sum(unblockedCountsMeans>5),sum(unblockedCountsMeans>10),sum(unblockedCountsMeans>50),sum(unblockedCountsMeans>100),sum(unblockedCountsMeans>500)))

#blocked: collapse technical replicates (by just taking mean of counts)
blockedCounts<- as.data.frame(counts[,9:16])
blockedCounts$ACollapsed<- rowMeans(blockedCounts[,1:3])
blockedCounts$BCollapsed<- rowMeans(blockedCounts[,4:6])
blockedCounts<-blockedCounts[,c(9,10,7,8)]

#get mean counts across all 4 samples
blockedCountsMeans<- as.data.frame((rowMeans(blockedCounts)))

#number that pass thresholds:
blockedPassThresholds<- data.frame(threshold=c(1,5,10,50,100,500), countsPassingThreshold=c(sum(blockedCountsMeans>1),sum(blockedCountsMeans>5),sum(blockedCountsMeans>10),sum(blockedCountsMeans>50),sum(blockedCountsMeans>100),sum(blockedCountsMeans>500)))

#combine into one, adding in an extra column for blocker status
unblockedPassThresholds$status<- rep("unblocked",6)
blockedPassThresholds$status<- rep("blocked",6)
passThreshold<-rbind(unblockedPassThresholds,blockedPassThresholds)

#plot
p<- ggplot(passThreshold, aes(x=threshold, y=countsPassingThreshold, color=status))+
             geom_line(size=2)+
  geom_point(size=4)+
  scale_color_manual(values=c("black", "red"))+
  ggtitle("Number of miRNAs passing a count threshold")+
  theme_classic()
p
```

#doing the above, but not taking the mean. still collapse technical, but keep a/b/c/d info. Add in a 4th column in the final table with sample type. Then subtract the #passing threshold for a/b/c/d individually from blocker vs no blocker, then graph. 


```{r}
#Set desired thresholds
thresholds<-c(1,3,5,10,25,50,75,100,250,500,600,700)
NumberInEachGroup<- length(thresholds)*5 #used for plotting later

#Get counts passing threshold for unblocked samples, then transform data- 4 columns: ID (A/b/c/d), threshold, #of species that passed that threshold, and blocker status
unblockedPassThreshold<- as.data.frame(apply(unblockedCounts, 2, 
                                             CountsPassingThreshold, thresholds))
rownames(unblockedPassThreshold)<- thresholds
colnames(unblockedPassThreshold)<- c("A", "B", "C", "D")
#add 4th column: mean of all samples
unblockedPassThreshold$Mean<- rowMeans(unblockedPassThreshold)
unblockedPassThreshold<- as.data.frame(t(unblockedPassThreshold))
unblockedPassThreshold$ID<- rownames(unblockedPassThreshold)
#reformat- melt
unblockedPassThreshold<- melt(unblockedPassThreshold)
colnames(unblockedPassThreshold)<- c("ID", "threshold","speciesPassingThreshold")
unblockedPassThreshold$status<- rep("unblocked", NumberInEachGroup)
unblockedPassThreshold$threshold<- as.integer(as.character((unblockedPassThreshold$threshold)))

#Get counts passing threshold for blocked samples, then transform data- 4 columns: ID (A/b/c/d), threshold, #of species that passed that threshold, and blocker status
blockedPassThreshold<- as.data.frame(apply(blockedCounts, 2, 
                                             CountsPassingThreshold, thresholds))
rownames(blockedPassThreshold)<- thresholds
colnames(blockedPassThreshold)<- c("A", "B", "C", "D")
#add 4th column: mean of all samples
blockedPassThreshold$Mean<- rowMeans(blockedPassThreshold)
blockedPassThreshold<- as.data.frame(t(blockedPassThreshold))
blockedPassThreshold$ID<- rownames(blockedPassThreshold)
#reformat- melt
blockedPassThreshold<- melt(blockedPassThreshold)
colnames(blockedPassThreshold)<- c("ID", "threshold","speciesPassingThreshold")
blockedPassThreshold$status<- rep("blocked", NumberInEachGroup)
blockedPassThreshold$threshold<- as.integer(as.character((blockedPassThreshold$threshold)))

#combine into one
passThreshold<-rbind(unblockedPassThreshold,blockedPassThreshold)

#reformat so that you have 4 columns: ID, threshold, #species passing in noblocker, #species passing in blocker
PT<-data.frame(passThreshold$ID[1:NumberInEachGroup],
               passThreshold$threshold[1:NumberInEachGroup],
               passThreshold$speciesPassingThreshold[1:NumberInEachGroup],
               passThreshold$speciesPassingThreshold[(NumberInEachGroup+1):nrow(passThreshold)])
colnames(PT)<- c("ID", "threshold", "SpeciesPT_noblocker", "SpeciesPT_blocker")

#now need to subtract each- i.e., sample A at threshold 1 with no blocker vs with blocker. get those values for all, then plot. 
df<- as_tibble(PT)
df<-df%>% group_by(ID) %>% mutate(BlockerMinusNoBlocker=SpeciesPT_blocker-SpeciesPT_noblocker)
df<- as.data.frame(df)
df$Mean<- rep(c(rep("Sample",4),"Mean"),length(thresholds))
df$Mean<- as.factor(df$Mean)

Samplecolors<- c("goldenrod1","violetred3", "skyblue3", "darkolivegreen4")

#plot
p<- ggplot(df, aes(x=threshold, y=BlockerMinusNoBlocker, color=ID))+
  geom_line(aes(linetype = Mean, group=ID, size=Mean))+
  scale_linetype_manual(values= c("solid", "dashed"),guide=FALSE)+
  scale_size_manual(values=c("Sample"=1.2,"Mean"=2), guide=FALSE)+
  #geom_point(size=1.3)+
  xlab("Count threshold")+
  ylab("miRNA species passing threshold (blocked minus unblocked)")+
  scale_y_continuous(breaks=seq(-50, 400, 50), limits=c(0,350))+
  scale_x_continuous(breaks=seq(0, max(df$threshold), 100))+
  scale_color_manual(values=c("goldenrod1","violetred3", "skyblue3", "darkolivegreen4", "black"))+
  theme_classic()+
  guides(color = guide_legend(override.aes = list(size = 2)))+
  theme(axis.text.y   = element_text(size=12, colour = "black", face= "bold"),
        axis.title.y  = element_text(size=16, colour = "black", face= "bold"),
        axis.text.x   = element_text(size=12, colour = "black", face= "bold"),
        axis.title.x  = element_text(size=16, colour = "black", face= "bold"),
        panel.border = element_rect(colour = "black", fill=NA, size=2),
        legend.title=element_blank(),
        legend.position = c(.9,.85),
        legend.background = element_rect(fill=NA, size=1, linetype="solid", color="black"),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"),
        legend.text=element_text(size=14, face="bold"))
p

#Exported as Figure5A_SpeciesPassingThreshold_032520

```



#Figure 5B- Differential expression
General idea: compare the number of significantly DE species between A and B in both blocked and unblocked samples.

Two analyses:

1) The 3 A.0 samples vs the 3 B.0 samples
2) The 3 A.1 samples vs the 3 B.1 samples


Analysis #1: DESeq2 using A0 vs B0 - here, technical replicates are NOT collapsed
```{r}
#rename rows, removing input info
A0_B0<- counts[,c(1:6)]
colnames(A0_B0)<- c("A.0.1", "A.0.2", "A.0.3", "B.0.1", "B.0.2", "B.0.3")


#create metadata
sampleGroup<- c(rep("A", 3), rep("B", 3))
sample<- colnames(A0_B0)
meta<- data.frame(sample, sampleGroup)

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=A0_B0, 
                             colData=meta, 
                             design=~sampleGroup)

#Run DESeq2
dds<- DESeq(dds)

#Extrac results
res<- na.omit(results(dds))
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,]) #89
nrow(resOrdered[resOrdered$padj<0.01,]) #60

#Save for later
noblocker_AvB<- resOrdered
```

DESeq2 using A1 vs B1- here, technical replicates are NOT collapsed
```{r}
#rename rows, removing input info
A1_B1<- counts[,c(9:14)]
colnames(A1_B1)<- c("A.1.1", "A.1.2", "A.1.3", "B.1.1", "B.1.2", "B.1.3")

#create metadata
sampleGroup<- c(rep("A", 3), rep("B", 3))
sample<- colnames(A1_B1)
meta<- data.frame(sample, sampleGroup)

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=A1_B1, 
                             colData=meta, 
                             design=~sampleGroup)

#Run DESeq2
dds<- DESeq(dds)

#Extract results
res<- na.omit(results(dds))
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,]) #119
nrow(resOrdered[resOrdered$padj<0.01,]) #75

#Save for later
blocker_AvB<- resOrdered
```


```{r}
#So that's a....
(119-89)/119
#25% increase in the number of DE genes!!
```

#Figure 5B
```{r}
#Get list of sig miRs at given sig levels 
SigLevels<- c(0.1, 0.05, 0.01)
blocked_Sig<-lapply(SigLevels, GetSigSpecies, res=blocker_AvB)
unblocked_Sig<-lapply(SigLevels, GetSigSpecies, res=noblocker_AvB)

#Get number of sig miRs at given sig levels
blocked_NSig<- unlist(lapply(blocked_Sig, nrow))
unblocked_NSig<- unlist(lapply(unblocked_Sig, nrow))

#Merge into one df for plotting
df<- data.frame(Sig=as.factor(SigLevels), Blocked=blocked_NSig, Unblocked=unblocked_NSig)
df<- melt(df)
df$Order<- c(rep(2,3), rep(1,3))

#plot:
p<- ggplot(df, aes(x=reorder(Sig, -value), y=value, fill=variable, group=Order))+
        geom_bar(position="dodge", stat="identity")+
        scale_fill_manual(values=c(Blocked="firebrick3", Unblocked="turquoise3"))+
        theme_classic()+
        xlab("Signficance threshold (p<)")+
        ylab("Number of Significant miRs")+
        theme(axis.title.x=element_text(size=16, face="bold"),
              axis.title.y=element_text(size=16, face="bold"),
              axis.text.x= element_text(size=14, face="bold", color="black"),
              axis.text.y=element_text(size=14, face="bold", color="black"),
              legend.title = element_blank(),
              legend.text = element_text(size=14, face="bold"),
              legend.background = element_rect(fill=NA, size=1, 
                                  linetype="solid", color="black"),
              legend.position = c(.72,.9),
              panel.border = element_rect(colour = "black", fill=NA, size=2))


```

#Figure 5C
```{r}
#Get miRs that are specific to blocked/unblocked at each sig level
blocked_unique<-list()
unblocked_unique<- list()

for (i in 1:length(blocked_Sig)){
        blocked<- blocked_Sig[[i]]
        unblocked<- unblocked_Sig[[i]]
        blocked_unique[[i]]<- blocked[!(rownames(blocked) %in% rownames(unblocked)),]
        unblocked_unique[[i]]<- unblocked[!(rownames(unblocked)%in% rownames(blocked)),]
}


#get raw counts for the blocked/unblocked specific miRs
blocked_unique_rawCounts<- lapply(blocked_unique, function(x){
        counts[rownames(counts) %in% rownames(x),]
})

unblocked_unique_rawCounts<- lapply(unblocked_unique, function(x){
        counts[rownames(counts) %in% rownames(x),]
})

#Get median of the mean raw count for blocked/unblocked specific miRs
blocked_unique_MedianofMeanRawCounts<- unlist(lapply(blocked_unique_rawCounts, function(x){
        median(rowMeans(x))
}))

unblocked_unique_MedianofMeanRawCounts<- unlist(lapply(unblocked_unique_rawCounts, function(x){
        median(rowMeans(x))
}))

#Get median rowMeans for all DE miRs common between blocked and unblocked
CommonDE_RawCounts<-list()
for (i in 1:length(blocked_Sig)){
        blocked<- blocked_Sig[[i]]
        unblocked<- unblocked_Sig[[i]]
        CommonDE<- rownames(blocked[rownames(blocked) %in% rownames(unblocked),])
        CommonDE_RawCounts[[i]]<- counts[rownames(counts) %in% CommonDE,]
}

CommonDE_MedianofMeanRawCounts<- unlist(lapply(CommonDE_RawCounts, function(x){
        median(rowMeans(x))
}))

#merge into one df for plotting
df<- data.frame(Sig=as.factor(SigLevels), 
                Blocked=blocked_unique_MedianofMeanRawCounts, 
                Unblocked=unblocked_unique_MedianofMeanRawCounts,
                CommonDE=CommonDE_MedianofMeanRawCounts)
                

df<-melt(df)
df$Order<- c(rep(3,3), rep(2,3), rep(1,3))

#plot:
p<- ggplot(df, aes(x=reorder(Sig, -value), y=value, fill=variable, group=Order))+
        geom_bar(position="dodge", stat="identity")+
        scale_fill_manual(values=c(Blocked="firebrick3", Unblocked="turquoise3", CommonDE="grey"))+
        theme_classic()+
        xlab("Signficance threshold (p<)")+
        ylab("Median Raw counts for DE miRs \n specific to blocked or unblocked DE analysis")+
        theme(axis.title.x=element_text(size=16, face="bold"),
              axis.title.y=element_text(size=16, face="bold"),
              axis.text.x= element_text(size=14, face="bold", color="black"),
              axis.text.y=element_text(size=14, face="bold", color="black"),
              legend.title = element_blank(),
              legend.text = element_text(size=10, face="bold"),
              legend.background = element_rect(fill=NA, size=1, 
                                  linetype="solid", color="black"),
              legend.position = c(.88,.9),
              panel.border = element_rect(colour = "black", fill=NA, size=2))

```


Plot these differences in DE miRNAs for 3 different cutoffs. Three graphs:

1) Number of DE species for blocked/unblocked at 3 sig cutoffs
2) Number of DE Species UNIQUE to that group for blocked/unblocked at 3 sig cutoffs
3) Median raw read count for these UNIQUE DE species at 3 sig cutoffs

```{r}
UnblockedSig<-noblocker_AvB
BlockedSig<- blocker_AvB
SigLevel<-0.01
RawCounts<-counts

#Get data frame of sig miRNAs for blocked and unblocked at given significance level
  NBSig<-UnblockedSig[UnblockedSig$padj<SigLevel,]
  BSig<- BlockedSig[BlockedSig$padj<SigLevel,]
  
  #Get DE miRNA specific to blocked and unblocked
  DEInUB<- rownames(as.data.frame(BSig)) %in% rownames(as.data.frame(NBSig))
  DEInB<- rownames(as.data.frame(NBSig)) %in% rownames(as.data.frame(BSig))
  OnlyBlockedDE<- BSig[!DEInUB,]
  OnlyUnBlockedDE<- BSig[!DEInB,]
  
  #Get raw counts for these miRNAs
  counts_onlyblockedDE<- RawCounts[rownames(RawCounts) %in% rownames(OnlyBlockedDE),]
  counts_onlyblockedDE<- counts_onlyblockedDE[,c(1:6,9:14)]
  counts_onlyblockedDE<-counts_onlyblockedDE[order(counts_onlyblockedDE$A.0, decreasing=TRUE),]
  
  counts_onlyunblockedDE<- RawCounts[rownames(RawCounts) %in% rownames(OnlyUnBlockedDE),]
  counts_onlyunblockedDE<- counts_onlyunblockedDE[,c(1:6,9:14)]
  counts_onlyunblockedDE<-counts_onlyunblockedDE[order(counts_onlyunblockedDE$A.1, decreasing=TRUE),]
  
  return(list(counts_onlyblockedDE, counts_onlyunblockedDE))
```


```{r}
#Merge blocker and unblocked lists, for each sig level.Only keeping species that are DE in both analyses
Sig.01<- merge(as.data.frame(blocker_AvB[blocker_AvB$padj<0.01,]),
               as.data.frame(noblocker_AvB[noblocker_AvB$padj<0.01,]), by=0)
Sig.01Counts<- counts[rownames(counts)%in% Sig.01$Row.names,]

Sig.05<- merge(as.data.frame(blocker_AvB[blocker_AvB$padj<0.05,]),
               as.data.frame(noblocker_AvB[noblocker_AvB$padj<0.05,]), by=0)
Sig.05Counts<- counts[rownames(counts)%in% Sig.05$Row.names,]

Sig.1<- merge(as.data.frame(blocker_AvB[blocker_AvB$padj<0.1,]),
               as.data.frame(noblocker_AvB[noblocker_AvB$padj<0.1,]), by=0)
Sig.1Counts<- counts[rownames(counts)%in% Sig.1$Row.names,]

#padj<0.01
Median_All_0.01<-median(rowMeans(Sig.01Counts))
Sig_Counts_Blocked_0.01<- GetCounts_UniqueDEmiRNA(blocker_AvB, noblocker_AvB, 0.01, counts)[[1]]
Sig_Counts_UnBlocked_0.01<- GetCounts_UniqueDEmiRNA(blocker_AvB, noblocker_AvB, 0.01, counts)[[2]]
Median_Blocked_0.01<- median(rowMeans(Sig_Counts_Blocked_0.01[,1:6,])) #14.33333
Median_UnBlocked_0.01<-median(rowMeans(Sig_Counts_UnBlocked_0.01[,7:12,])) #651
NSig_Blocked_0.01<-nrow(Sig_Counts_Blocked_0.01) #30
NSig_UnBlocked_0.01<-nrow(Sig_Counts_UnBlocked_0.01) #16

#padj<0.05
Median_All_0.05<-median(rowMeans(Sig.05Counts))
Sig_Counts_Blocked_0.05<- GetCounts_UniqueDEmiRNA(blocker_AvB, noblocker_AvB, 0.05, counts)[[1]]
Sig_Counts_UnBlocked_0.05<- GetCounts_UniqueDEmiRNA(blocker_AvB, noblocker_AvB, 0.05, counts)[[2]]
Median_Blocked_0.05<-median(rowMeans(Sig_Counts_Blocked_0.05[,1:6,])) #12.33333
Median_UnBlocked_0.05<-median(rowMeans(Sig_Counts_UnBlocked_0.05[,7:12,])) #249.4167
NSig_Blocked_0.05<-nrow(Sig_Counts_Blocked_0.05) #56
NSig_UnBlocked_0.05<-nrow(Sig_Counts_UnBlocked_0.05) #30

#padj<0.1
Median_All_0.1<-median(rowMeans(Sig.1Counts))
Sig_Counts_Blocked_0.1<- GetCounts_UniqueDEmiRNA(blocker_AvB, noblocker_AvB, 0.1, counts)[[1]]
Sig_Counts_UnBlocked_0.1<- GetCounts_UniqueDEmiRNA(blocker_AvB, noblocker_AvB, 0.1, counts)[[2]]
Median_Blocked_0.1<-median(rowMeans(Sig_Counts_Blocked_0.1[,1:6,])) #8
Median_UnBlocked_0.1<-median(rowMeans(Sig_Counts_UnBlocked_0.1[,7:12,])) #254.6667
NSig_Blocked_0.1<-nrow(Sig_Counts_Blocked_0.1) #77
NSig_UnBlocked_0.1<-nrow(Sig_Counts_UnBlocked_0.1) #33

#Dataframe summarizing n sig miRNAs + median counts for sig miRNAs specific to blocked/unblocked, for the 3 different significance levels
Sig_BvNB<- data.frame(SignificanceLevel= as.factor(c(rep(0.01, 2), rep(0.05 ,2), rep(0.1, 2))),
                      SignificantSpecies=c(nrow(blocker_AvB[blocker_AvB$padj<0.01,]), 
                                           nrow(noblocker_AvB[noblocker_AvB$padj<0.01,]),
                                           nrow(blocker_AvB[blocker_AvB$padj<0.05,]),
                                           nrow(noblocker_AvB[noblocker_AvB$padj<0.05,]),
                                           nrow(blocker_AvB[blocker_AvB$padj<0.1,]),
                                           nrow(noblocker_AvB[noblocker_AvB$padj<0.1,])),
                      UniqueSignificantSpecies=c(NSig_Blocked_0.01, NSig_UnBlocked_0.01,
                                           NSig_Blocked_0.05, NSig_UnBlocked_0.05,
                                           NSig_Blocked_0.1, NSig_UnBlocked_0.1),
                      MedianUniqueCounts= c(Median_Blocked_0.01, Median_UnBlocked_0.01,
                                            Median_Blocked_0.05, Median_UnBlocked_0.05,
                                            Median_Blocked_0.1, Median_UnBlocked_0.1),
                      BlockerStatus= c(rep(c("Blocked", "Unblocked"), 3)))

BvNB_MedianCounts<- data.frame(SignificanceLevel=as.factor(c(rep(0.01,3), rep(0.05,3), rep(0.1,3))), 
                               MedianCounts=c(Median_Blocked_0.01, Median_UnBlocked_0.01,Median_All_0.01,
                                              Median_Blocked_0.05, Median_UnBlocked_0.05,Median_All_0.05,
                                              Median_Blocked_0.1, Median_UnBlocked_0.1,Median_All_0.1),
                               BlockerStatus= c(rep(c("Blocked", "Unblocked", "All"), 3)))
BvNB_MedianCounts$BlockerStatus<-factor(BvNB_MedianCounts$BlockerStatus, levels=c("Blocked", "Unblocked", "All"))

#Plot this data on four separate graphs
p_NSpecies<- ggplot(Sig_BvNB, aes(y=SignificantSpecies, x=SignificanceLevel, fill=BlockerStatus))+
  geom_bar(position="dodge", stat="identity")+
  theme_classic()+
  xlab("Significance Level Threshold")+
  ylab("Number of significantly DE species")+
  scale_fill_manual("legend", values=c("Blocked"="firebrick3", "Unblocked"="blue3"))+
  ylim(0,200)+
  theme(axis.text.y   = element_text(size=10, colour = "black"),
        axis.title.y  = element_text(size=12, colour = "black"),
        axis.text.x   = element_text(size=10, colour = "black"),
        axis.title.x  = element_text(size=12, colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=2),
        legend.title=element_blank(),
        legend.position = c(.85,.96),
        legend.background = element_rect(fill=NA, size=1, linetype="solid", color="black"),
        axis.text.x.bottom  =element_text(size=15),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"))
  
p_UniqueNSpecies<-  ggplot(Sig_BvNB, aes(y=UniqueSignificantSpecies, x=SignificanceLevel, fill=BlockerStatus))+
  geom_bar(position="dodge", stat="identity")+
  theme_classic()+
  xlab("Significance Level Threshold")+
  ylab("Number of significantly DE species (for species unique to specific sample group)")+
  scale_fill_manual("legend", values=c("Blocked"="firebrick3", "Unblocked"="blue3"))+
  ylim(0,200)+
  theme(axis.text.y   = element_text(size=10, colour = "black"),
        axis.title.y  = element_text(size=12, colour = "black"),
        axis.text.x   = element_text(size=10, colour = "black"),
        axis.title.x  = element_text(size=12, colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=2),
        legend.title=element_blank(),
        legend.position = c(.85,.96),
        legend.background = element_rect(fill=NA, size=1, linetype="solid", color="black"),
        axis.text.x.bottom  =element_text(size=15),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm")) 

p_MedianCounts<- ggplot(Sig_BvNB, aes(y=MedianUniqueCounts, x=SignificanceLevel, fill=BlockerStatus))+
  geom_bar(position="dodge", stat="identity")+
  theme_classic()+
  xlab("Significance Level Threshold")+
  ylab("Median raw counts (for species unique to specific sample group)")+
  scale_fill_manual("legend", values=c("Blocked"="firebrick3", "Unblocked"="blue3"))+
  theme(axis.text.y   = element_text(size=10, colour = "black"),
        axis.title.y  = element_text(size=12, colour = "black"),
        axis.text.x   = element_text(size=10, colour = "black"),
        axis.title.x  = element_text(size=12, colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=2),
        legend.title=element_blank(),
        legend.position = c(.85,.96),
        legend.background = element_rect(fill=NA, size=1, linetype="solid", color="black"),
        axis.text.x.bottom  =element_text(size=15),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"))

p_MedianCounts<- ggplot(BvNB_MedianCounts, aes(y=MedianCounts, x=SignificanceLevel, fill=BlockerStatus))+
  geom_bar(position="dodge", stat="identity")+
  theme_classic()+
  xlab("Significance Level Threshold")+
  ylab("Median raw counts (for species unique to specific sample group)")+
  scale_fill_manual("legend", values=c("Blocked"="firebrick3", "Unblocked"="blue3", "All"="gray"))+
  theme(axis.text.y   = element_text(size=10, colour = "black"),
        axis.title.y  = element_text(size=12, colour = "black"),
        axis.text.x   = element_text(size=10, colour = "black"),
        axis.title.x  = element_text(size=12, colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=2),
        legend.title=element_blank(),
        legend.position = c(.85,.9),
        legend.background = element_rect(fill=NA, size=1, linetype="solid", color="black"),
        axis.text.x.bottom  =element_text(size=15),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"))
  

#grid.arrange(p_NSpecies+ggtitle(""),p_UniqueNSpecies, p_MedianCounts+ggtitle(""), nrow=1)

#ggarrange(p_NSpecies, p_UniqueNSpecies, p_MedianCounts, nrow=1)

```
