---
title: "Blocker Project: Differential Expression using DESeq2"
output: html_notebook
---

#Using DESeq2 to perform differential expression on miRNASeq data, comparing blocked vs no blocked groups


Import libraries:
```{r}
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(pheatmap))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(metaseqR))
```


#Two different input sources: 

1) count matrix from BaseSpace (miRDeep2, no mismatches, isomirs and mature together) --> Downsampled to lowest read count
    a) Isomirs and mature
    b) Mature only
2) count matrix generated from miRDeep2 at the command line- mature only --> Downsampled to lowest read count


#For #1b and #2 count matrices, 2 different DESeq2 analyses performed:

1) Standard- Pairwise, technical replicates collapsed, all mature species included
2) Excluding targets- same as standard, but only including NON TARGET mature species
    a) excluding just the actual targets
    b) excluding the actual targets + members of the same family

Matrix #1 used initially, but after determining that a mismatch of 1 should be allowed (i.e., miRDeep2 default settings), the miRDeep2 analysis was rerun at the command line (matrix #2) with default miRDeep2 parameters. Both analyses used the same input bam file, generated from BaseSpace (splice-aware STAR alignmenet)


#Notes on DESeq2 Analysis:

-Since there are 4 sample types (A/B/C/D) each with at least one blocked/unblocked pair, pairwise DESeq2 analysis is performed

-Samples A and B have 3 technical replicates, for both blocked and unblocked groups. These technical replicates are collapsed using an internal DESeq2 function. These technical replicates have varying RNA input amounts into library construction, but this was determined to have no effect on the resulting libraries (See ___.Rmd). This is additionally confirmed by PCA clustering here.

Read in data:
```{r}
wd<- "//Cifs2/rcdata$/Channing_miRNA_Blocker_Test/"
BaseSpace_countsInput<- read.delim(paste(wd, "TGC_Data/mirna-blocker_bs_mature-isomirs_raw.txt", sep =""))
CL_countsInput<- "placeholder"

#Reformat both count matrices
BaseSpace_counts<- BaseSpace_countsInput
rownames(BaseSpace_counts)<- BaseSpace_counts$ID
BaseSpace_counts<- BaseSpace_counts[,c(4:(ncol(BaseSpace_counts)))]
#3,455 mature/isomir species

BaseSpace_counts<- BaseSpace_countsInput
rownames(BaseSpace_counts)<- BaseSpace_counts$ID
BaseSpace_counts<- BaseSpace_counts[,c(4:(ncol(BaseSpace_counts)))]

```

########################################################################################
##Using Matrix #1 (generated by BaseSpace, no mismatches, mature and isomirs included):#
########################################################################################

Two analyses with this count matrix performed:

a) Using both mature and isomir data. Isomirs are collapsed into one (e.g., if there are multiple isomirs for hsa-mir-92a, these counts are all added together)
b) Using only mature data

First: Downsample Matrix #1 to the lowest number of reads. This downsampled matrix will be used for all downstream analyses.

```{r}
#Using package metaseqr, downsample counts to lowest total count- here, that's sample C.1, with 7,970,014 isomir and mature counts
#set seed so results are replicable
BaseSpace_counts_downsample<- downsample.counts(BaseSpace_counts, seed=42)
BaseSpace_counts<- BaseSpace_counts_downsample
```

Write to csv
```{r}
write.csv(BaseSpace_counts_downsample, paste(wd, "OtherResults/DownSampledCounts_BaseSpaceMatureIsomir_metaseqR_121119.csv", sep =""))
```

Analysis A: Mature and Isomir

Prep data for DESeq2, create metadata and DESeq2 object, then run DESeq2:
```{r}
#rename rows, removing input info
colnames(BaseSpace_counts)<- c(rep("A.0", 3), rep("B.0", 3), "C.0", "D.0", rep("A.1", 3), rep("B.1", 3), "C.1", "D.1")

#Create metadata:

#set blocker status
sampleGroup<- as.factor(c(rep("NoBlocker", 8), rep("Blocker", 8)))
#this column is used to collapse technical replicates
sample<- colnames(BaseSpace_counts)
#this column is to control for sample to sample differences
patient<- as.factor(c(rep("A", 3), rep("B", 3), "C", "D", rep("A", 3), rep("B", 3), "C", "D"))
meta<- data.frame(sample, sampleGroup, patient)

#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later (i.e., pairwise analysis)
dds<- DESeqDataSetFromMatrix(countData=BaseSpace_counts, 
                             colData=meta, 
                             design=~sampleGroup)

#Collapse replicates- using internal DESeq2 function
ddsColl<- collapseReplicates(dds, dds$sample)

#Create model that takes into account sample differences
ddsMF<- ddsColl
design(ddsMF)<- formula(~patient + sampleGroup)

#run DESeq2
dds<- DESeq(ddsMF)
```

Visualize sample differences using PCA

Note: this function uses the dds object, but no information on sample group is input into the analysis (i.e., unsupervised clustering)
```{r}
#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(ddsMF, blind=FALSE)
PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ylim(-12,12)+
  ggtitle("PCA: Blocker vs no Blocker, Mature + Isomirs")+
  geom_text(aes(label=name),hjust=.5, vjust=-1.2, color="black")
  
p
```

Samples do cluster by blocker status as we would expect, but they also cluster strongly by sample type- A/B are more similar to one another than C/D. 

Is there any possibility this is due to collapsing of technical replicates? If we didn't collapse technical replicates, would we see all 3 A.1/3 B.1, etc still all clustering together?
Testing this..
```{r}
TechNotCollapsed<-BaseSpace_counts

#Rename samples
colnames(TechNotCollapsed)<- c(paste("A.0.", 1:3,sep=""), paste("B.0.", 1:3,sep=""), "C.0", "D.0", paste("A.1.", 1:3,sep=""), paste("B.1.", 1:3,sep=""), "C.1", "D.1")

#Create metadata:

#set blocker status
sampleGroup<- as.factor(c(rep("NoBlocker", 8), rep("Blocker", 8)))
#this column is to control for sample to sample differences
patient<- as.factor(c(paste("A", 1:3,sep=""), paste("B", 1:3,sep=""), "C", "D", paste("A", 1:3,sep=""), paste("B", 1:3,sep=""), "C", "D"))
meta<- data.frame(sampleGroup, patient)

#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later (i.e., pairwise analysis)
dds_TNC<- DESeqDataSetFromMatrix(countData=TechNotCollapsed, 
                             colData=meta, 
                             design=~sampleGroup)


#Create model that takes into account sample differences
ddsMF_TNC<- dds_TNC
design(ddsMF_TNC)<- formula(~patient + sampleGroup)

#run DESeq2
dds_TNC<- DESeq(ddsMF_TNC)

#Normalized (variance stabilized transformation), then run PCA
vsd_TNC <- varianceStabilizingTransformation(ddsMF_TNC, blind=FALSE)
PCA_TNC<-plotPCA(vsd_TNC, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA_TNC, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ylim(-15,15)+
  ggtitle("PCA: Blocker vs no Blocker, Technical Replicates not Collapsed: Mature + Isomirs")+
  geom_text(aes(label=name),hjust=.5, vjust=-1.2, color="black")
  
p

```
So the answer to that question is yes- samples still cluster in a similar pattern, even if you don't collapse by technical replicates. A/B and C/D really are more similar to one another.


#Back to using the dds object with technical replicates collapsed:


Differential expression results:
```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
nrow(resOrdered[resOrdered$padj<0.01,])
#54 DE miRNAs- padj <0.05
#31 DE miRNAs- padj <0.01

rownames(resOrdered[resOrdered$padj<0.01,])
```


Visualize DE results with heatmap:
```{r}
#normalize results
rld<- rlog(dds)

#Select top 75 sig DE miRNAs
topgenes<- rownames(resOrdered[1:75,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

#Heatmap settings
subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)

#Create metadata- to be used to color code to show which group the sample belongs to
metadata<- data.frame(Sample_Group=PCA$sampleGroup, row.names=PCA$name)

#plot heatmap
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)),show_rownames = FALSE, annotation_col = metadata, border_color= NA)
```

Blocked/unblocked samples cluster together, as expected. Samples A/B and C/D also cluster together, as seen in the PCA results


#DESeq2 Analysis B, still using count matrix #1

Count matirx #1 = BaseSpace miRDeep2 count matrix, no mismatches
Analysis B = only mature miRNAs- isomirs excluded

```{r}
#Select just mature miRNAs. let7 needs to be searched for separately (different nomenclature)
BaseSpace_MatureCounts<- BaseSpace_counts[grep("R", rownames(BaseSpace_counts)),]
let7<- BaseSpace_counts[grep("let", rownames(BaseSpace_counts)),]
BaseSpace_MatureCounts<- rbind(BaseSpace_MatureCounts, let7)
BaseSpace_MC<- BaseSpace_MatureCounts
nrow(BaseSpace_MC)
#2,599 mature miRNA species
```

Prep data for DESeq2, create metadata and DESeq2 object, then run DESeq2:
```{r}
#rename rows, removing input info
colnames(BaseSpace_MC)<- c(rep("A.0", 3), rep("B.0", 3), "C.0", "D.0", rep("A.1", 3), rep("B.1", 3), "C.1", "D.1")

#Create metadata:

#set blocker status
sampleGroup<- as.factor(c(rep("NoBlocker", 8), rep("Blocker", 8)))
#this column is used to collapse technical replicates
sample<- colnames(BaseSpace_MC)
#this column is to control for sample to sample differences
patient<- as.factor(c(rep("A", 3), rep("B", 3), "C", "D", rep("A", 3), rep("B", 3), "C", "D"))
meta<- data.frame(sample, sampleGroup, patient)

#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later (i.e., pairwise analysis)
dds<- DESeqDataSetFromMatrix(countData=BaseSpace_MC, 
                             colData=meta, 
                             design=~sampleGroup)

#Collapse replicates- using internal DESeq2 function
ddsColl<- collapseReplicates(dds, dds$sample)

#Create model that takes into account sample differences
ddsMF<- ddsColl
design(ddsMF)<- formula(~patient + sampleGroup)

#run DESeq2
dds<- DESeq(ddsMF)

```

Visualize sample differences using PCA
```{r}
#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(ddsMF, blind=FALSE)
PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ylim(-12,12)+
  ggtitle("PCA: Blocker vs no Blocker, Mature Only")+
  geom_text(aes(label=name),hjust=.5, vjust=-1.2, color="black")
  
p
```

As with mature/isomir data, sample still cluster by blocker status and sample type in similar clustering pattern


Differential expression results:
```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.01,])
nrow(resOrdered[resOrdered$padj<0.05,])
#20 DE miRNAs- padj <0.05
#15 DE miRNAs- padj <0.01

rownames(resOrdered[resOrdered$padj<0.05,])
```

Visualize DE results with heatmap:
```{r}
#normalize results
rld<- rlog(dds)

#Select top 25 sig DE miRNAs- fewer here than for mature/isomir analysis because there are a lot fewer sig DE miRNAs
topgenes<- rownames(resOrdered[1:25,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

#Heatmap settings
subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)

#Create metadata- to be used to color code to show which group the sample belongs to
metadata<- data.frame(Sample_Group=PCA$sampleGroup, row.names=PCA$name)

#plot heatmap
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)),show_rownames = FALSE, annotation_col = metadata, border_color= NA)
```
Samples still cluster by blocker status and sample type, like the mature/isomir dataset. Note: if you increase the number of species plotted here, you no longer see samples clustering by blocker status. This is likely not a concern, since only a dozen or so species are actually sig DE


Write results to file:
```{r}
#write to file 
write.csv(resOrdered, paste(wd, "DESeq2Results/DESeq2Results_MatureOnlyBaseSpaceCountMatrix_Pairwise_121119.csv", sep = "" ))
```


#DESeq2 Analysis 2: removing targeted miRNAs

Two variations
    a) excluding just the actual targets
    b) excluding the actual targets + members of the same family

#Variation A (only actual targets removed) using count matrix #1b (from BaseSpace, mature species only)

```{r}
#Get list of all targets
ActualTargets<- c('hsa-miR-486-5p','hsa-miR-92a-3p','hsa-miR-451a')

#subset counts (mature only) by non-target miRNAs
BS_MC_AT<- BaseSpace_MC[!(rownames(BaseSpace_MC) %in% ActualTargets),]
```

Create dds object, run DESeq2
```{r}
#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later (i.e., pairwise analysis)
dds<- DESeqDataSetFromMatrix(countData=BS_MC_AT, 
                             colData=meta, 
                             design=~sampleGroup)

#Collapse replicates- using internal DESeq2 function
ddsColl<- collapseReplicates(dds, dds$sample)

#Create model that takes into account sample differences
ddsMF<- ddsColl
design(ddsMF)<- formula(~patient + sampleGroup)

#run DESeq2
dds<- DESeq(ddsMF)
```


Visualize sample differences using PCA
```{r}
#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(ddsMF, blind=FALSE)
PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ylim(-12,12)+
  ggtitle("PCA: Blocker vs no Blocker, Mature Only")+
  geom_text(aes(label=name),hjust=.5, vjust=-1.2, color="black")
  
p
```

When you remove the targets from analysis, samples cluster completely by SAMPLE TYPE! Exactly what we would expect. Those 3 targets are causing the blocker status clustering we see in all other PCA plots shown previously here. Suggests that there are no huge off target effects caused by the blocker that cause large expression differences between blocked/unblocked


Extract DESeq2 results and visualize with heatmap
```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.01,])
nrow(resOrdered[resOrdered$padj<0.05,])
#15 DE miRNAs- padj <0.05
#12 DE miRNAs- padj <0.01

rownames(resOrdered[resOrdered$padj<0.05,])

#normalize results
rld<- rlog(dds)

#Select top 25 sig DE miRNAs- fewer here than for mature/isomir analysis because there are a lot fewer sig DE miRNAs
topgenes<- rownames(resOrdered[1:75,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

#Heatmap settings
subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)

#Create metadata- to be used to color code to show which group the sample belongs to
metadata<- data.frame(Sample_Group=PCA$sampleGroup, row.names=PCA$name)

#plot heatmap
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)),show_rownames = FALSE, annotation_col = metadata, border_color= NA)
```

In heatmap of DE results, samples cluster by SAMPLE and not BLOCKER STATUS, as we would expect if there are no huge off target effects caused by the blockers. 

Note: If you cluster using <50 species, A and B still cluster by sample type, but C/D cluster by blocker status (C.0/D.0 and C.1/D.1). Only when you take a broader look at expression, like above, do they cluster by sample type. This begins to suggest that there may be a few species that do have some differences in expression between blocked/unblocked. This will be looked at more closely in ____.Rmd




#Variation B (actual targets + species in same family removed) using count matrix #1b (from BaseSpace, mature species only)

Select all species with 92, 451, or 486 in name. Also select 2 hsa-mir-25 species (same family as 92a)
```{r}
Target_92<- rownames(BS_MC_AT[grep("miR-92", rownames(BS_MC_AT))[1:4],])
Target_451<- rownames(BS_MC_AT[grep("miR-451", rownames(BS_MC_AT))[1],])
Target_486<- rownames(BS_MC_AT[grep("miR-486", rownames(BS_MC_AT)),])
Target_25<- rownames(BS_MC_AT[grep("miR-25", rownames(BS_MC_AT)),])
TargetFamilies<- c(Target_451,Target_486,Target_92, Target_25)

#Using BaseSpace Mature counts matrix that already has the actual targets removed, remove the 8 additional species that are in the same family as the 3 actual targets
BS_MC_NoTF<- BS_MC_AT[!(rownames(BS_MC_AT)%in%TargetFamilies),]
```



Create dds object, run DESeq2, visualize with PCA
```{r}
#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later (i.e., pairwise analysis)
dds<- DESeqDataSetFromMatrix(countData=BS_MC_NoTF, 
                             colData=meta, 
                             design=~sampleGroup)

#Collapse replicates- using internal DESeq2 function
ddsColl<- collapseReplicates(dds, dds$sample)

#Create model that takes into account sample differences
ddsMF<- ddsColl
design(ddsMF)<- formula(~patient + sampleGroup)

#run DESeq2
dds<- DESeq(ddsMF)

#Visualize sample differences using PCA

#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(ddsMF, blind=FALSE)
PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ylim(-12,12)+
  ggtitle("PCA: Blocker vs no Blocker, Mature Only, Targets + Target Families removed")+
  geom_text(aes(label=name),hjust=.5, vjust=-1.2, color="black")
  
p
```

Very similar to previous analysis (just actual targets removed), samples still cluster by SAMPLE TYPE not by blocker status

Extract DESeq2 results and visualize with heatmap
```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
nrow(resOrdered[resOrdered$padj<0.01,])
#14 DE miRNAs- padj <0.05
#10 DE miRNAs- padj <0.01

rownames(resOrdered[resOrdered$padj<0.05,])

#normalize results
rld<- rlog(dds)

#Select top 25 sig DE miRNAs- fewer here than for mature/isomir analysis because there are a lot fewer sig DE miRNAs
topgenes<- rownames(resOrdered[1:15,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

#Heatmap settings
subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)

#Create metadata- to be used to color code to show which group the sample belongs to
metadata<- data.frame(Sample_Group=PCA$sampleGroup, row.names=PCA$name)

#plot heatmap
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)),show_rownames = FALSE, annotation_col = metadata, border_color= NA)
```

Samples still cluster by sample type, not blocker status!

IMPORTANT: when you remove not just the actual targets, but also the actual targets + species from the same family (as above), samples cluster by sample type EVEN WHEN YOU SELECT JUST A FEW (as low as 15) SPECIES TO CLUSTER BY

Compare this to when you remove just the actual targets: you have to select at least 50 miRNAs to see all samples clustering by sample type rather than blocker status

This just further increases our confidence that there are very few off target effects caused by blockers (other than blocking members of the same family, which can't really be fully avoided)


