---
title: "Effect of blockers on targeted miRNAs and closely related species"
output: html_notebook
---

#With addition of blockers, the percent of reads that align to targets decreases

Import libraries
```{r}
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(pheatmap))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(gridExtra))
```

Functions
```{r}
#For plottting expression levels of targets
PlotExpressionLevels<- function(PercentMapToTargets, ylab){
  p<- ggplot(PercentMapToTargets, aes(y=PercentMapToTarget, x=Sample, fill=BlockerStatus))+
  geom_bar(stat="identity", position="dodge")+ 
  theme_classic()+
  ylab(ylab)+
  xlab("")+
  scale_fill_manual("legend", values=c("Blocked"="firebrick3", "Unblocked"="blue3"))+
  ylim (0,100)+
  theme(axis.text.y   = element_text(size=10, colour = "black"),
        axis.title.y  = element_text(size=12),
        axis.title.x  = element_text(size=14),
        panel.border = element_rect(colour = "black", fill=NA, size=2),
        legend.title=element_blank(),
        legend.position = c(.1,.94),
        legend.background = element_rect(fill=NA, size=1, linetype="solid", color="black"),
        axis.text.x.bottom  =element_text(size=15),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"))
  return(p)
}

#For plotting expression levels of targets, broken down by target
PlotIndividualExpressionLevels<- function(PercentMapToTargets, ylab){
  p<- ggplot(PercentMapToTargets, aes(y=PercentMapToTarget, x=ID, fill=BlockerStatus))+
  geom_bar(stat="identity", position="dodge")+ 
  theme_classic()+
  ylab(ylab)+
  xlab("")+
  ylim (0,100)+
  scale_fill_manual("legend", values=c("Blocked"="firebrick3", "Unblocked"="blue3"))+
  facet_wrap(~species,nrow=1)+
  theme(axis.text.y   = element_text(size=10, colour = "black"),
        axis.title.y  = element_text(size=12),
        axis.title.x  = element_text(size=14),
        panel.border = element_rect(colour = "black", fill=NA, size=2), 
                       strip.background = element_blank(),
        legend.title=element_blank(),
        legend.position = c(.1,.92),
        legend.background = element_rect(fill=NA, size=1, linetype="solid", color="black"),
        axis.text.x.bottom  =element_text(color="black", size=15),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"),
        strip.text = element_text(face="bold"))
  return(p)
  
}

#For plotting without collapsing technical replicates

PlotUncollapsedExpression<- function(PercentMapToTargets, ylab){
  p<-ggplot(PercentMapToTargets, aes(y=PercentMapToTarget, x=ID, fill=BlockerStatus))+
  geom_bar(stat="identity", position="dodge")+ 
  theme_classic()+
  ylab(ylab)+
  xlab("")+
  ylim (0,100)+
  scale_fill_manual("legend", values=c("Blocked"="firebrick3", "Unblocked"="blue3"))+
  theme(axis.text.y   = element_text(size=10, colour = "black"),
        axis.title.y  = element_text( size=12),
        axis.title.x  = element_text(size=14),
        panel.border = element_rect(colour = "black", fill=NA, size=2),
        legend.title=element_blank(),
        legend.position = c(.1,.92),
        legend.background = element_rect(fill=NA, size=1, linetype="solid", color="black"),
        legend.key.size = unit(.4, "cm"),
        axis.text.x.bottom  =element_text(color= "black",size=15),
        legend.spacing.y = unit(0, "mm"),
        legend.spacing.x = unit(0, "mm"))
  return(p)
}

```

#BaseSpace Data
################
Read in data. Data from basespace mirdeep2 analysis (no mismatches allowed for mature, isomirs included) and then downsampled (see Final_DifferentialExpression_BlockerProject.Rmd)
```{r}
wd<- "//Cifs2/rcdata$/Channing_miRNA_Blocker_Test/"
BaseSpaceDownSampled_Counts<- read.csv(paste(wd, "OtherResults/DownSampledCounts_BaseSpaceMatureIsomir_metaseqR_121119.csv", sep =""))
rownames(BaseSpaceDownSampled_Counts)<- BaseSpaceDownSampled_Counts$X
BaseSpaceDownSampled_Counts<- BaseSpaceDownSampled_Counts[,-1]
```

#Percent of reads mapping to targets in samples without blocker vs with blocker: Collapsing technical replicates

First, collapse technical replicates- average counts across technical replicates for A and B samples
```{r}
#Get mean of A/B samples (Collapsing technical replicates)
BS_CollapsedCounts<- BaseSpaceDownSampled_Counts
BS_CollapsedCounts$A.0_Collapsed<- as.integer(rowMeans(BaseSpaceDownSampled_Counts[,1:3]))
BS_CollapsedCounts$B.0_Collapsed<- as.integer(rowMeans(BaseSpaceDownSampled_Counts[,4:6]))
BS_CollapsedCounts$A.1_Collapsed<- as.integer(rowMeans(BaseSpaceDownSampled_Counts[,9:11]))
BS_CollapsedCounts$B.1_Collapsed<- as.integer(rowMeans(BaseSpaceDownSampled_Counts[,12:14]))
BS_CollapsedCounts<- BS_CollapsedCounts[,c(17,18,7,8,19,20,15,16)]

colnames(BS_CollapsedCounts)<- c("A.0", "B.0", "C.0", "D.0", "A.1", "B.1", "C.1", "D.1")
```


Next, get number of reads mapping to any of the targets and divide by the total number of reads
```{r}
#Get list of all targets (and very closely related miRNAs)
Target_92<- rownames(BaseSpaceDownSampled_Counts[grep("miR-92", rownames(BaseSpaceDownSampled_Counts), ignore.case=TRUE),])[1:8]
Target_451<- rownames(BaseSpaceDownSampled_Counts[grep("miR-451", rownames(BaseSpaceDownSampled_Counts), ignore.case=TRUE),])[c(1,4,9,10)]
Target_486<- rownames(BaseSpaceDownSampled_Counts[grep("miR-486", rownames(BaseSpaceDownSampled_Counts), ignore.case=TRUE),])
Target_25<- rownames(BaseSpaceDownSampled_Counts[grep("miR-25", rownames(BaseSpaceDownSampled_Counts), ignore.case=TRUE),])
TargetFamilies<- c(Target_451,Target_486,Target_92, Target_25)

#get total counts for these targets in all samples
targetCounts<- BS_CollapsedCounts[rownames(BS_CollapsedCounts)%in% TargetFamilies,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(BS_CollapsedCounts) *100)
PercentMapToTargets$ID<- rownames(PercentMapToTargets)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("Unblocked", 4), rep("Blocked", 4))
PercentMapToTargets$Sample<-rep(c("A", "B", "C", "D"),2)
```



Finally, visualize this change for all 4 samples from blocker to no blocker
```{r}
p<- PlotExpressionLevels(PercentMapToTargets,ylab="Percent of reads mapping to target miRNAs or species in target families (isomir matches allowed)")
```

#Repeating the above graph, but with only mature species

```{r}
#Select mature species
BaseSpace_MatureCounts<- BS_CollapsedCounts[grep("R", rownames(BS_CollapsedCounts)),]
let7<- BS_CollapsedCounts[grep("let", rownames(BS_CollapsedCounts)),]
BaseSpace_MatureCounts<- rbind(BaseSpace_MatureCounts, let7)
BaseSpace_MC<- BaseSpace_MatureCounts

#get total counts for these targets in all samples
targetCounts<- BaseSpace_MC[rownames(BaseSpace_MC)%in% TargetFamilies,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(BaseSpace_MC) *100)
PercentMapToTargets$ID<- rownames(PercentMapToTargets)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("Unblocked", 4), rep("Blocked", 4))
PercentMapToTargets$Sample<-rep(c("A", "B", "C", "D"),2)

#plot
p<- PlotExpressionLevels(PercentMapToTargets,"Percent of reads mapping to target miRNAs or species in target families (exact matches only)" )
```

#Mature Only, only looking at actual targets
```{r}
#get total counts for these targets in all samples
targetCounts<- BaseSpace_MC[rownames(BaseSpace_MC)%in% ActualTargets,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(BaseSpace_MC) *100)
PercentMapToTargets$ID<- rownames(PercentMapToTargets)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("Unblocked", 4), rep("Blocked", 4))
PercentMapToTargets$Sample<-rep(c("A", "B", "C", "D"),2)

#Plot
p<- PlotExpressionLevels(PercentMapToTargets, "Percent of reads mapping to target miRNAs (exact matches only)")
```

#Repeating the above graphs, but not collapsing technical replicates

Mature and Isomirs
```{r}
#get total counts for these targets in all samples
targetCounts<- BaseSpaceDownSampled_Counts[rownames(BaseSpaceDownSampled_Counts)%in% TargetFamilies,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(BaseSpaceDownSampled_Counts) *100)
PercentMapToTargets$ID<- rownames(PercentMapToTargets)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("Unblocked", 8), rep("Blocked", 8))

#Rename samples to low/medium/high RNA input
PercentMapToTargets$ID<- rep(c("A.1", "A.2", "A.3", "B.1", "B.2", "B.3", "C", "D"),2) 

#plot
p<- PlotUncollapsedExpression(PercentMapToTargets, "Percent of reads mapping to target miRNAs or species in target families (isomirs allowed)")
```

Only mature species:
```{r}
#Select Mature
BaseSpace_MatureCounts<- BaseSpaceDownSampled_Counts[grep("R", rownames(BaseSpaceDownSampled_Counts)),]
let7<- BaseSpaceDownSampled_Counts[grep("let", rownames(BaseSpaceDownSampled_Counts)),]
BaseSpace_MatureCounts<- rbind(BaseSpace_MatureCounts, let7)
BaseSpace_MC<- BaseSpace_MatureCounts

#get total counts for these targets in all samples
targetCounts<- BaseSpace_MC[rownames(BaseSpace_MC)%in% TargetFamilies,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(BaseSpace_MC) *100)
PercentMapToTargets$ID<- rownames(PercentMapToTargets)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("Unblocked", 8), rep("Blocked", 8))

#Rename samples 
PercentMapToTargets$ID<- rep(c("A.1", "A.2", "A.3", "B.1", "B.2", "B.3", "C", "D"),2) 


#plot
p<- PlotUncollapsedExpression(PercentMapToTargets, "Percent of reads mapping to target miRNAs or species in target families (exact matches only)")
```


#Expression levels broken down by each individual target

mature only --> Get reads mapped to each of the three targets in blocked vs unblocked

```{r}
#Select mature from collapsed data set
BaseSpace_MatureCounts<- BS_CollapsedCounts[grep("R", rownames(BS_CollapsedCounts)),]
let7<- BS_CollapsedCounts[grep("let", rownames(BS_CollapsedCounts)),]
BaseSpace_MatureCounts<- rbind(BaseSpace_MatureCounts, let7)
BaseSpace_MC<- BaseSpace_MatureCounts

#get number of reads mapping to 3 targets
mir451<- BaseSpace_MC[rownames(BaseSpace_MC) == "hsa-miR-451a",]
mir486<- BaseSpace_MC[rownames(BaseSpace_MC) == "hsa-miR-486-5p",]
mir92<- BaseSpace_MC[rownames(BaseSpace_MC) == "hsa-miR-92a-3p",]


#Get percentage of total counts that map to 3 targets. Add column for target information
PercentMapToTargets_451<-as.data.frame(t(as.data.frame(mir451/colSums(BaseSpace_MC) *100)))
PercentMapToTargets_451$species<- "hsa-miR-451a"
colnames(PercentMapToTargets_451)<- c("PercentMap", "species")
PercentMapToTargets_486<-as.data.frame(t(as.data.frame(mir486/colSums(BaseSpace_MC) *100)))
PercentMapToTargets_486$species<- "hsa-miR-486-5p"
colnames(PercentMapToTargets_486)<- c("PercentMap", "species")
PercentMapToTargets_92<-as.data.frame(t(as.data.frame(mir92/colSums(BaseSpace_MC) *100)))
PercentMapToTargets_92$species<- "hsa-miR-92a-3p"
colnames(PercentMapToTargets_92)<- c("PercentMap", "species")

#Combine 3 target dfs into one
PercentMapToTargets<- rbind(PercentMapToTargets_451,PercentMapToTargets_486,PercentMapToTargets_92)
PercentMapToTargets$ID<- rep(c("A", "B", "C", "D"),2)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "species", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("Unblocked", 4), rep("Blocked", 4))

#plot
p<- PlotIndividualExpressionLevels(PercentMapToTargets, "Percent of reads mapping to each target (exact matches only)")
```

Overall: mir-486 is blocked very well, mir-92 is blocked pretty well, and mir-451a is hardly even expressed

However- this is for the mature species only. Blockers could also be affecting isomirs

Repeating the above graph, but including isomirs:

```{r}
#get number of reads mapping to 3 targets
mir451_Both<- BS_CollapsedCounts[rownames(BS_CollapsedCounts) %in% c("hsa-miR-451a", "hsa-mir-451a"),]
mir451<- as.data.frame(colSums(mir451_Both))

mir486_Both<-BS_CollapsedCounts[rownames(BS_CollapsedCounts) %in% c("hsa-miR-486-5p", "hsa-mir-486-1","hsa-mir-486-2"),]
mir486<- as.data.frame(colSums(mir486_Both))

mir92_Both<- BS_CollapsedCounts[rownames(BS_CollapsedCounts) %in% c("hsa-miR-92a-3p", "hsa-mir-92a-1","hsa-mir-92a-2"),]
mir92<- as.data.frame(colSums(mir92_Both))


#Get percentage of total counts that map to 3 targets. Add column for target information
PercentMapToTargets_451<-as.data.frame(mir451/colSums(BS_CollapsedCounts) *100)
PercentMapToTargets_451$species<- "hsa-miR-451a"
colnames(PercentMapToTargets_451)<- c("PercentMap", "species")
PercentMapToTargets_486<-as.data.frame(mir486/colSums(BS_CollapsedCounts) *100)
PercentMapToTargets_486$species<- "hsa-miR-486-5p"
colnames(PercentMapToTargets_486)<- c("PercentMap", "species")
PercentMapToTargets_92<-as.data.frame(mir92/colSums(BS_CollapsedCounts) *100)
PercentMapToTargets_92$species<- "hsa-miR-92a-3p"
colnames(PercentMapToTargets_92)<- c("PercentMap", "species")

#Combine 3 target dfs into one
PercentMapToTargets<- rbind(PercentMapToTargets_451,PercentMapToTargets_486,PercentMapToTargets_92)
PercentMapToTargets$ID<- rep(c("A", "B", "C", "D"),2)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "species", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("Unblocked", 4), rep("Blocked", 4))

#Plot
p<- PlotIndividualExpressionLevels(PercentMapToTargets, "Percent of reads mapping to each target (isomirs allowed)")
```

Blockers don't seem to be quite so effective on isomirs as they are on mature, but we still see a large drop in target reads with the use of blockers. 

Can also see here tha mir-451a is expressed- it's just only expressed as an isomir primarily. "Isomir" for this dataset, however, can mean a single nucleotide change, even at the very end of the sequence.



#Command Line miRDeep2 Analysis

Repeating the above analysis, but using the count matrix generated by miRDeep2 at the command line, with default settings (1 mismatch)

Only mature miRNA are reported- no isomirs

These mature miRNA allow for 1 mismatch (or more in the 3'/5' end), compared to the previous count matrix from basespace that allows for ZERO mismatches

#Read in data
```{r}
wd<- "//Cifs2/rcdata$/Channing_miRNA_Blocker_Test/BlockerProject/BlockerProject/"
CL_countsDownsampledInput<- read.csv(paste(wd, "RawData/Downsampled_miRNA_CLmiRDeep2Results_AllSamples_DefaultmiRDeep2Settings.csv", sep =""))
rownames(CL_countsDownsampledInput)<- CL_countsDownsampledInput$X
CL_counts<- CL_countsDownsampledInput[,-1]

#reorder columns correctly
CL_counts<- CL_counts[,c(1,3,5,7,9,11,13,15,2,4,6,8,10,12,14,16)]
```


```{r}
#Get mean of A/B samples (Collapsing technical replicates)
CL_CollapsedCounts<- CL_counts
CL_CollapsedCounts$A.0_Collapsed<- as.integer(rowMeans(CL_counts[,1:3]))
CL_CollapsedCounts$B.0_Collapsed<- as.integer(rowMeans(CL_counts[,4:6]))
CL_CollapsedCounts$A.1_Collapsed<- as.integer(rowMeans(CL_counts[,9:11]))
CL_CollapsedCounts$B.1_Collapsed<- as.integer(rowMeans(CL_counts[,12:14]))
CL_CollapsedCounts<- CL_CollapsedCounts[,c(17,18,7,8,19,20,15,16)]

colnames(CL_CollapsedCounts)<- c("A.0", "B.0", "C.0", "D.0", "A.1", "B.1", "C.1", "D.1")
```

Next, get number of reads mapping to any of the targets and divide by the total number of reads
```{r}
#Get list of all targets (and very closely related miRNAs)
Target_92<- rownames(CL_CollapsedCounts[grep("miR-92", rownames(CL_CollapsedCounts))[5:9],])
Target_451<- rownames(CL_CollapsedCounts[grep("miR-451", rownames(CL_CollapsedCounts))[11:12],])
Target_486<- rownames(CL_CollapsedCounts[grep("miR-486", rownames(CL_CollapsedCounts)),])
Target_25<- rownames(CL_CollapsedCounts[grep("miR-25", rownames(CL_CollapsedCounts)),])
TargetFamilies<- c(Target_451,Target_486,Target_92, Target_25)

#get total counts for these targets in all samples
targetCounts<- CL_CollapsedCounts[rownames(CL_CollapsedCounts)%in% TargetFamilies,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(CL_CollapsedCounts) *100)
PercentMapToTargets$Sample<- rep(c("A", "B", "C", "D"),2)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "Sample")
PercentMapToTargets$BlockerStatus<- c(rep("Unblocked", 4), rep("Blocked", 4))
PercentMapToTargets
```


Finally, visualize this change for all 4 samples from blocker to no blocker
```{r}
p<-PlotExpressionLevels(PercentMapToTargets, "Percent of reads mapping to target miRNAs or miRNAs in target families (1 mismatch allowed)")
```



#Repeating the above graph, but only with mapping to actual targets

```{r}
#get total counts for these targets in all samples
ActualTargets<- c("hsa-miR-451a","hsa-miR-486-5p", "hsa-miR-92a-3p")
targetCounts<- CL_CollapsedCounts[rownames(CL_CollapsedCounts)%in% ActualTargets,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(CL_CollapsedCounts) *100)
PercentMapToTargets$Sample<- rep(c("A", "B", "C", "D"),2)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "Sample")
PercentMapToTargets$BlockerStatus<- c(rep("Unblocked", 4), rep("Blocked", 4))

#plot
p<- PlotExpressionLevels(PercentMapToTargets,"Percent of reads mapping to target miRNAs (1 mismatch allowed)" )
```

#Repeating the above graph, but not collapsing technical replicates
```{r}
#get total counts for these targets in all samples
targetCounts<- CL_counts[rownames(CL_counts)%in% ActualTargets,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(CL_counts) *100)
PercentMapToTargets$ID<- rownames(PercentMapToTargets)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("Unblocked", 8), rep("Blocked", 8))

#Rename samples 
PercentMapToTargets$ID<- rep(c("A1", "A2", "A3", "B1", "B2", "B3", "C", "D"),2) 


#plot
p<- PlotUncollapsedExpression(PercentMapToTargets,"Percent of reads mapping to target miRNAs ( 1 mismatch allowed)")
```



#Percent mapping broken down by individual target
```{r}
#get number of reads mapping to 3 targets
mir451_Both<- CL_CollapsedCounts[rownames(CL_CollapsedCounts) %in% c("hsa-miR-451a", "hsa-mir-451a"),]
mir451<- as.data.frame(colSums(mir451_Both))

mir486_Both<-CL_CollapsedCounts[rownames(CL_CollapsedCounts) %in% c("hsa-miR-486-5p", "hsa-mir-486-1","hsa-mir-486-2"),]
mir486<- as.data.frame(colSums(mir486_Both))

mir92_Both<- CL_CollapsedCounts[rownames(CL_CollapsedCounts) %in% c("hsa-miR-92a-3p", "hsa-mir-92a-1","hsa-mir-92a-2"),]
mir92<- as.data.frame(colSums(mir92_Both))


#Get percentage of total counts that map to 3 targets. Add column for target information
PercentMapToTargets_451<-as.data.frame(mir451/colSums(CL_CollapsedCounts) *100)
PercentMapToTargets_451$species<- "hsa-miR-451a"
colnames(PercentMapToTargets_451)<- c("PercentMap", "species")
PercentMapToTargets_486<-as.data.frame(mir486/colSums(CL_CollapsedCounts) *100)
PercentMapToTargets_486$species<- "hsa-miR-486-5p"
colnames(PercentMapToTargets_486)<- c("PercentMap", "species")
PercentMapToTargets_92<-as.data.frame(mir92/colSums(CL_CollapsedCounts) *100)
PercentMapToTargets_92$species<- "hsa-miR-92a-3p"
colnames(PercentMapToTargets_92)<- c("PercentMap", "species")

#Combine 3 target dfs into one
PercentMapToTargets<- rbind(PercentMapToTargets_451,PercentMapToTargets_486,PercentMapToTargets_92)
PercentMapToTargets$ID<- rep(c("A", "B", "C", "D"),2)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "species", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("Unblocked", 4), rep("Blocked", 4))

#plot
p<- PlotIndividualExpressionLevels(PercentMapToTargets, "Percent of reads mapping to target miRNA (1 mismatch allowed)")
```

Overall- using perfect or default settings in miRDeep2 gives a similar pattern- blockers are shown as effective in both. However, it looks like blockers are a bit more effective against PERFECT matches to its targets. You can see this in the better blocking of mature vs isomirs in basespace data


Testing:
```{r}
library(Biostrings)

isomirs<- read.csv("//Cifs2/rcdata$/Channing_miRNA_Blocker_Test/ChanningBlockerPilot_RProject/RawCounts/miRNA_isomirs_downloadedSeparately.csv")
isomirs$Sequence<- as.character(isomirs$Sequence)
test<- na.omit(as.data.frame(isomirs %>% group_by(ID) %>% summarise(mean=mean(nchar(Sequence)), sd=sd(nchar(Sequence)), min=min(nchar(Sequence)), max=max(nchar(Sequence))) %>% mutate(minDiff= 22-min) %>% mutate(maxDiff=max-22)))

test486<- isomirs[grep("let-7a-1", isomirs$ID),]

dna <-readDNAStringSet(test486$Sequence)
stringDist(dna, method="hamming")
```

