---
title: "R Notebook"
output: html_notebook
---

#Large cohort: differential expression of paired subset. Plotting heatmaps and Log2FC Blocked vs Unblocked.

Import libraries
```{r}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(metaseqR))
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(pheatmap))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(gsubfn))
windowsFonts("Arial" = windowsFont("Arial"))
```


Functions
```{r}
#For plotting log2 CPM values nicely
CPMPlot<- function(CPM, Sample){
  #Get subset of CPM for labeling points- points that are both sig DE and are higher in unblocked than   blocked (beneath the slope line)
  CPM$HigherInUnblocked<- (CPM$Blocker-CPM$NoBlocker)<0
  SubsetToLabel<- CPM[CPM$HigherInUnblocked=="TRUE" &CPM$Color!="Not DE",]
  
  #Plot
  p<- ggplot(CPM, aes(x=NoBlocker, y=Blocker, color=Color))+
  geom_point(size=3)+
  geom_abline(slope=1, color="red")+
  scale_colour_manual(values=c("red","black", "blue"))+
  theme_classic()+
  xlim(0,6)+
  ylim(0,6)+
  xlab(paste("Log of Unblocked Counts Per Million"))+
  ylab(paste("Log of Blocked Counts Per Million"))+
  geom_text_repel(data=SubsetToLabel, aes(x=NoBlocker, y=Blocker, label=rownames(SubsetToLabel)), color="black",  size=5, fontface="bold")+
  ggtitle(Sample)+
  theme(axis.text.y   = element_text(size=14, colour = "black", face="bold"),
        axis.title.y  = element_text(size=14, colour = "black", face="bold"),
        axis.title.x  = element_text(size=14, colour = "black", face="bold"),
        axis.text.x  = element_text(size=14, colour = "black", face="bold"),
        legend.background = element_rect(fill=NA, size=1, 
                                         linetype="solid", color="black"),
        legend.title=element_blank(),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"),
        legend.position = c(.15,.892),
        legend.text = element_text(size=18, face="bold"),
        panel.border = element_rect(colour = "black", fill=NA, size=2),
        plot.title=element_text(size=14, face="bold"))
  return(p)
  
}

#Plot PCA nicely
MyPCA<- function(dds, metadata, PC1Variance, PC2Variance,miny, maxy, minx, maxx) {
  #Normalized (variance stabilized transformation), then run PCA
  vsd <- varianceStabilizingTransformation(dds, blind=TRUE)
  PCA<-plotPCA(vsd, intgroup="Batch",returnData=TRUE)
  PCA$blocker<- metadata$blocker
  
  #plot
  p<- ggplot(PCA, aes(x=PC1, y=PC2, color=Batch, shape=blocker))+
  geom_point(size=4)+
  #scale_color_manual(values=Groupcolors)+
  scale_shape_manual(values=c(16,18))+
  theme_classic()+
  ylim(miny,maxy)+
  xlim(minx,maxx)+
  ylab(paste("PC2: ", PC2Variance, "% Variance", sep =""))+
  xlab(paste("PC1: ", PC1Variance, "% Variance", sep =""))+
  #geom_text(aes(label=Sample),hjust=.5, size= 5,vjust=-1.2, color="black")+
  theme(axis.text.y   = element_text(size=12, colour = "black", face="bold"),
        axis.title.y  = element_text(size=14, colour = "black", face="bold"),
        axis.title.x  = element_text(size=14, colour = "black", face="bold"),
        axis.text.x  = element_text(size=12, colour = "black", face="bold"),
        legend.background = element_rect(fill=NA, size=1, 
                                         linetype="solid", color="black"),
        legend.title=element_blank(),
        legend.text = element_text(size=14, face="bold"),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"),
        legend.position = c(.9,.85),
        panel.border = element_rect(colour = "black", fill=NA, size=2)
        )
  return(p)
}

```
#Overall Steps:

1: Read in paired count matrixes and subject IDs of paired samples, merge counts together and collapse mature/isomiRs. Filter based on CPM.
2: Create DESeq2 object, run pair-wise DE on all species and plot heatmap
3: Create DESeq2 object, run pair-wise DE on NON-TARGET species (miR-486/miR-451a/miR-92a/miR-92b/miR-25 removed priot to DE)
4: Get off targets (From targets included DESeq2, step 2), plot. Export potential off-targets to csv.

```{r}
#read in paired data, blocked and unblocked
wd<-"~/BlockerProject/"
unblocked_paired<-read.csv(paste(wd, "LargeCohort_PairedSubset_89Samples_Unblocked.csv", sep=""),row.names="X") 
blocked_paired<-read.csv(paste(wd, "LargeCohort_PairedSubset_89Samples_Blocked.csv", sep=""),row.names="X") 

#Read in paired samples- From BH. Add column for new sampleID- "sampleIDppm_Blocked/Unblocked". Needed since some sample IDs are replicated between blocked/unblocked. 
pairedSamples<-read.csv(paste(wd, "PairedSamples_LargeCohort_FromBH.csv", sep=""), row.names = "UniqueID")
pairedSamples<-pairedSamples[,-1]

#merge counts together
counts_paired<- merge(unblocked_paired, blocked_paired, by=0)
rownames(counts_paired)<-counts_paired$Row.names
counts_paired<-counts_paired[,-1]
counts_paired<-counts_paired[,order(colnames(counts_paired))]

#Collapse counts together for species in same family- -3p/5p/-1/-2
counts_paired$geneIDs<- sub("^([^-]*-[^-]*-[^-]*).*", "\\1", rownames(counts_paired))
counts_paired$geneIDs<- gsub("mir", "miR", counts_paired$geneIDs)
counts_collapsed_paired<-counts_paired %>% group_by(geneIDs) %>% summarise_each(funs(sum)) %>% as.data.frame
rownames(counts_collapsed_paired)<-counts_collapsed_paired$geneIDs
counts_collapsed_paired<- counts_collapsed_paired[,2:ncol(counts_collapsed_paired)]

#Filter to remove any species without >1 CPM in >=75% of samples =="detected"
CPM<-(counts_collapsed_paired/colSums(counts_collapsed_paired))*1000000
CPM_Above1<- CPM[apply(CPM, 1, function(x) {sum(x>=1)})>(ncol(CPM)*.75),]
counts_CPMPass<-counts_collapsed_paired[rownames(counts_collapsed_paired) %in% rownames(CPM_Above1),]
```



#2: Create DESeq2 object, run pair-wise DE on all species

```{r}
#Read in batch info, merge with other metadata
batches<-read.csv(paste(wd, "PairedSample_Batches.csv", sep=""))
rownames(batches)<-paste(batches$Paired.samples, batches$BlockerStatus,sep="_")
pairedSamples<-merge(batches, pairedSamples, by=0)
pairedSamples<-pairedSamples[,c(1,5,4,6)]

#create DESeq2 object- just based on samplegroup (blockerstatus)
#will add samples to model later (i.e., pairwise analysis)
pairedSamples$blocker<-as.factor(pairedSamples$blocker)
pairedSamples$Batch<-as.factor(pairedSamples$Batch) #remove
pairedSamples$phenoId<-as.factor(pairedSamples$phenoId)

dds<- DESeqDataSetFromMatrix(countData=counts_CPMPass, 
                       colData=pairedSamples, 
                       design=~blocker)

vsd<-varianceStabilizingTransformation(dds)

dds$blocker<- relevel(dds$blocker, ref = "Unblocked")

#Create model that takes into account sample differences
ddsMF<- dds
design(ddsMF)<- formula(~phenoId + blocker)

#run DESeq2
dds<- DESeq(ddsMF)

#extract and order results  
res_paired<-na.omit(results(dds))
resOrdered_paired<-res_paired[order(res_paired$padj),]

nrow(resOrdered_paired[resOrdered_paired$padj<0.05,]) #270 (385 if unpaired, 170 if paired+CPM 75% filter)
nrow(resOrdered_paired[resOrdered_paired$padj<0.01,]) #244 (295 if unpaired, 154 if paired+CPM 75% filter)
nrow(resOrdered_paired[resOrdered_paired$padj<0.01 & abs(resOrdered_paired$log2FoldChange)>1,]) #138 (206 if unpaired, 69 if paired+CPM 75% filter)
```

Plot PCA- coloring by batch effects
```{r}
colors<-c("darkgoldenrod4", "darkgoldenrod2","darkcyan", "indianred4", "indianred1","lightblue4", "mediumpurple4",
          "grey77", "grey35","navy", "red4","darkolivegreen4", "orchid4")
p<-MyPCA(dds, pairedSamples, 29,13,-8,25,-20,25)+  scale_color_manual(values=colors)
```

Export results
```{r}
write.csv(resOrdered_paired, paste(wd, "PairedCohort_DESeq2Results_Collapsed.csv",sep=""))
```

Plot heatmap
```{r}
rld<-varianceStabilizingTransformation(dds)

MinSubtract=-3
MaxSubtract=2
nspecies=50
fontsize=12 

#Select top sig DE miRNAs
topgenes<- rownames(resOrdered_paired[1:nspecies,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

#Heatmap settings
mat_breaks<- seq(min(mat-MinSubtract), max(mat-MaxSubtract), length.out=75)

metadata<-pairedSamples[,c(1,2)]
rownames(metadata)<-metadata$Row.names

  #plot heatmap
  p<-pheatmap(mat, breaks = mat_breaks, 
            color =colorRampPalette( c("red", "black", "green"), space="rgb")(100),
            show_rownames = TRUE, show_colnames=FALSE,
            annotation = metadata,
            
            fontsize_row = fontsize,fontsize=12,treeheight_row=0,
            border_color= NA)
```



#Rerun pairwise DESeq2, removing 3 targets from analysis + removing 92b and 25 
Using count matrix that has been filtered to remove lowly expressed species (<50% of samples have >1 CPM)

```{r}
#Subset counts matrix, removing 3 targets
counts_collapsed_paired_RAT<- counts_CPMPass[!(rownames(counts_CPMPass) %in% c("hsa-miR-486",
                                                                                "hsa-miR-451a",
                                                                                "hsa-miR-92a",
                                                                                "hsa-miR-92b",
                                                                                "hsa-miR-25")),]

 
#create DESeq2 object- just based on samplegroup (blockerstatus)
#will add samples to model later (i.e., pairwise analysis)
dds<- DESeqDataSetFromMatrix(countData=counts_collapsed_paired_RAT, 
                       colData=pairedSamples, 
                       design=~blocker)
dds$blocker<- relevel(dds$blocker, ref = "Unblocked")

#Create model that takes into account sample differences
ddsMF<- dds
design(ddsMF)<- formula(~phenoId + blocker)

#run DESeq2
dds<- DESeq(ddsMF)

#extract and order results  
res_paired_RAT<-na.omit(results(dds))
resOrdered_paired_RAT<-res_paired_RAT[order(res_paired_RAT$padj),]


nrow(resOrdered_paired_RAT[resOrdered_paired_RAT$padj<0.05,]) #275 ( 216 if CPM filtered) 
nrow(resOrdered_paired_RAT[resOrdered_paired_RAT$padj<0.01,]) #243 ( 206 if CPM filtered)
nrow(resOrdered_paired_RAT[resOrdered_paired_RAT$padj<0.01 & abs(resOrdered_paired_RAT$log2FoldChange)>1,]) #142 ( 106 if CPM filtered)
```

Export results
```{r}
write.csv(resOrdered_paired_RAT, paste(wd, "PairedCohort_DESeq2Results_ActualTargetsRemoved.csv",sep=""))
```


Plot heatmap
```{r}
#rld<-varianceStabilizingTransformation(dds)

MinSubtract=0
MaxSubtract=2
nspecies=50
  #Select top sig DE miRNAs
  topgenes<- rownames(resOrdered_paired_RAT[1:nspecies,])
  #topgenes<-rownames(sig)
  mat<- assay(rld)[topgenes,]
  mat<- mat - rowMeans(mat)
  #colnames(mat)<-paste(metadata$Patient," (",metadata$sampleGroup, ")",sep="")
  
  #Heatmap settings
  mat_breaks<- seq(min(mat-MinSubtract), max(mat-MaxSubtract), length.out=75)

  #Create metadata- to be used to color code to show which group the sample belongs to
  #metadata<- data.frame(Group=pairedSamples$blocker, 
                       #row.names=rownames(pairedSam),
                        #Sample=metadata$Patient)
  metadata<-pairedSamples

  #plot heatmap
  p<-pheatmap(mat, breaks = mat_breaks, 
            color =colorRampPalette( c("red", "black", "green"), space="rgb")(100),
            show_rownames = TRUE, show_colnames=FALSE,
            annotation = metadata,
            
            fontsize_row = fontsize,fontsize=12,treeheight_row=0,
            border_color= NA)
```


#4: Get off targets (From targets included DESeq2, step 2), plot. Export potential off-targets to csv.
 
```{r}
#Get raw counts- only downsampled + isomirs collapsed. Separate into blocked/unblocked
counts<-counts_collapsed_paired
Unblocked_counts_collapsed_paired<-counts[,gsub(".*_", "", colnames(counts))=="Unblocked"]
Blocked_counts_collapsed_paired<-counts[,gsub(".*_", "", colnames(counts))=="Blocked"]

#Get sig DE genes- p<0.01 and baseMean>50
sig<-resOrdered_paired[resOrdered_paired$padj<0.01,]
sig<-sig[sig$baseMean>=50,]
sig<- sig[abs(sig$log2FoldChange)>=1,] #53, 101 if set at 0.5

#Add in CPM data for these miRNAs- across all 16 libraries
        #Convert to CPM
        counts_cpm<- (counts/colSums(counts)) *1000000
        
        #Get CPM for blocked and unblocked for sig miRs
        Unblocked_SigCPM<- as.data.frame(rowMeans(Unblocked_counts_collapsed_paired[rownames(Unblocked_counts_collapsed_paired) 
                                                                                    %in% rownames(sig),]))
        Blocked_SigCPM<- as.data.frame(rowMeans(Blocked_counts_collapsed_paired[rownames(Blocked_counts_collapsed_paired) 
                                                                                %in% rownames(sig),]))
        
        #Merge with DE data, reformat
        OT<- merge(as.data.frame(sig[,c(1,2,6)]), Unblocked_SigCPM, by=0)
        rownames(OT)<-OT$Row.names
        OT<- merge(OT, Blocked_SigCPM, by=0)
        OT<-OT[,-1]
        colnames(OT)<- c("Species", "baseMean", "log2FoldChange", "padj", "MeanCPMUnblockedCounts", "MeanCPMBlockedCounts")

#Get mean of CPM- used for plotting- and reformat species name
OT$MeanCPM<- rowMeans(OT[,5:6])
OT$Species<-gsub("hsa-", "", OT$Species)

#Annotate by whether species is a target miR
ActualTargets<- c('miR-486-5p','miR-92a-3p','miR-451a') #CL
ActualTargets<- c('miR-486','miR-92a','miR-451a') #BS
OT$Target<- OT$Species %in% ActualTargets

#Entire list of all species in target family
Target_92<- rownames(counts[grep("miR-92", rownames(counts)),])
Target_451<- rownames(counts[grep("miR-451", rownames(counts)),]) 
Target_486<- rownames(counts[grep("miR-486", rownames(counts)),]) 
Target_25<- rownames(counts[grep("miR-25", rownames(counts)),]) #CL or BS
TargetFamilies<- gsub("hsa-", "", c(Target_451,Target_486,Target_92, Target_25))

#Add family info in
OT$TargetFamily<- OT$Species %in% TargetFamilies
OT$Color<- as.factor(rowSums(OT[,c(8,9)]))
OT$Color<- gsub(0, "Non-Target miR", OT$Color)
OT$Color<- gsub(1, "miR in Target Family", OT$Color)
OT$Color<- gsub(2, "Target miR", OT$Color)


#set which miRs to be labeled on graph
SubsetToLabel<- OT[OT$Color %in% c("Target miR", "miR in Target Family"),]
SubsetToLabel$Species<- gsub("hsa-", "", SubsetToLabel$Species)
```

Write results to csv
```{r}
write.csv(OT, paste(wd,"OffTargetStats_PairedCohort_PairwiseDESeq2_TargetsIncluded_75CPMFilter.csv", sep=""))
```

Plot off target effects- same as in figure 3A
```{r}
#plot
p<-ggplot(OT,aes(x=log2FoldChange, y= log2(padj+1e-100), fill=Color))+
  geom_point(aes(size=MeanCPM), pch=21, color="black")+
  scale_size_continuous(range = c(1, 50), guide=FALSE)+
  #scale_x_continuous(limits=c(-6,6),n.breaks = 7)+
  #scale_y_continuous(limits = c(-0.008, .05))+
  theme_classic()+
  xlab("log2 foldchange of differential expression")+
  ylab("Log2(Adjusted p value)")+
  geom_vline(xintercept=0,linetype="dotted", color= "black", size=1.25)+
  guides(colour = guide_legend(override.aes = list(size=5)))+
  geom_text_repel(data=SubsetToLabel, aes(x=log2FoldChange, y=padj, 
                                          label=Species),
                                          size=3.5, fontface= "bold",
                                          nudge_y = 0.0035, point.padding = 1.7, 
                                          box.padding = 0.5)+
  scale_fill_manual(values=c("orange","black", "red"))+
  theme(axis.text.y   = element_text(size=12, face= "bold", colour = "black"),
        axis.title.y  = element_text(size=16, face= "bold", colour = "black"),
        axis.title.x  = element_text(size=16, face="bold", colour = "black"),
        axis.text.x  = element_text(size=12, face="bold", colour = "black"),
        legend.background = element_rect(fill=NA, size=1, 
                                         linetype="solid", color="black"),
        legend.title=element_blank(),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"),
        legend.position = c(.8,.93),
        panel.border = element_rect(colour = "black", fill=NA, size=2),
        legend.text = element_text(size=13, face="bold"))

p

```



#5: Plot log2 CPM values for blocked and unblocked datasets, coloring based on sig (from pairwise targets included analysis)
```{r}
#Convert to CPM
CPM<- counts_collapsed_paired/colSums(counts_collapsed_paired) *1000000

#Get mean CPM for blocked and unblocked for every species
blockedSamples<-pairedSamples[pairedSamples$blocker=="Blocked",]
blocked_CPM<-CPM[,colnames(CPM) %in% rownames(blockedSamples)]

unblockedSamples<-pairedSamples[pairedSamples$blocker=="Unblocked",]
unblocked_CPM<-CPM[,colnames(CPM) %in% rownames(unblockedSamples)]

CPM_Mean<-data.frame(Blocker=rowMeans(blocked_CPM), NoBlocker=rowMeans(unblocked_CPM))

#log2 of CPM
CPMLog<- log2(CPM_Mean+1) #+1 optional here

#Get miRNA info. Need: whether or not it was targeted + whether or not it was DE
#Add in target info- add column for whether or not the species was targeted
ActualTargets<- c('hsa-miR-486','hsa-miR-92a','hsa-miR-451a')
CPMLog$target<- rownames(CPMLog) %in% ActualTargets

#Add in DE info- using DESeq2 results from targets included analysis. Import data here.
CPMLog$Sig<- rownames(CPMLog) %in% rownames(sig)

#Add in third column, used to color points in graph. Add together target/sig columns. If it's a target and sig, it'll be 2, if it's just sig it'll be 1, if it's neither it'll be 0.
CPMLog$Color<- as.factor(rowSums(CPMLog[,3:4]))
CPMLog$Color<- gsub(1, "DE", CPMLog$Color)
CPMLog$Color<- gsub(0, "Not DE", CPMLog$Color)
CPMLog$Color<-gsub(2, "Target", CPMLog$Color)
CPMLog$Color<-factor(CPMLog$Color, levels=c("DE", "Not DE", "Target"))
CPMLog<-CPMLog[order(CPMLog$target, decreasing=TRUE),]

#plot
p<-CPMPlot(CPMLog, "Mean")+
        theme(axis.ticks = element_line(colour = "black", size =2), 
              axis.ticks.length = unit(2,"mm"))+
        xlim(0,20)+
        ylim(0,20)+
        theme(axis.title.x = element_text(size=20),
              axis.title.y = element_text(size=20))


```
