---
title: "R Notebook"
output: html_notebook
---

#Large Cohort: Paired and unpaired (full) Datasets: improvement in number of species passing count threshold


Import libraries
```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(metaseqR))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(pheatmap))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(gsubfn))
windowsFonts("Arial" = windowsFont("Arial"))
```

Functions
```{r}
#Function that takes a column of counts and returns the number of species that pass the given list of threhsolds
CountsPassingThreshold<- function(counts, threshold){
  Passthreshold<-c()
  for (i in 1:length(threshold)){
    Passthreshold[i]<- sum(counts>threshold[i])
    
  }
  return(Passthreshold)
}

Head<-function(x){x[1:10,1:10]}
```


#Overall steps:

1: Read in count matrixes- full cohort and paired cohort- and preprocess by collapsing together mature/isomiR
2: For full cohort: plot the number of species passing count thresholds
3: For paired cohort: plot the change in blocked/unblocked number of species passing count thresholds


#1: Read in count matrixes- full cohort and paired cohort- and preprocess by collapsing together mature/isomiR
```{r}
#unpaired, full dataset
unblocked_counts_Full<-read.csv(paste(wd, "UnblockedRawCounts_TechCollapsed.csv",sep=""),row.names="X") #unpaired full dataset
blocked_counts_Full<-read.csv(paste(wd, "BlockedRawCounts_TechCollapsed.csv",sep=""),row.names="X") #unpaired full dataset

#paired data, blocked and unblocked
unblocked_counts_paired<-read.csv(paste(wd, "LargeCohort_PairedSubset_89Samples_Unblocked.csv", sep=""),row.names="X") #paired subset
blocked_counts_paired<-read.csv(paste(wd, "LargeCohort_PairedSubset_89Samples_Blocked.csv", sep=""),row.names="X") #paired subset


#Collapse counts-remove everything after 3rd "-" (-1, -5p, etc), then merge and sum together based on new miR IDs

#Full Dataset:
        blocked_counts_Full$geneIDs<- sub("^([^-]*-[^-]*-[^-]*).*", "\\1", rownames(blocked_counts_Full))
        blocked_counts_Full$geneIDs<- gsub("mir", "miR", blocked_counts_Full$geneIDs)
        blocked_counts_Full_collapsed<-blocked_counts_Full %>% group_by(geneIDs) %>% summarise_each(funs(sum)) %>% as.data.frame
        rownames(blocked_counts_Full_collapsed)<-blocked_counts_Full_collapsed$geneIDs
        blocked_counts_Full_collapsed<- blocked_counts_Full_collapsed[,2:ncol(blocked_counts_Full_collapsed)]
        
        unblocked_counts_Full$geneIDs<- sub("^([^-]*-[^-]*-[^-]*).*", "\\1", rownames(unblocked_counts_Full))
        unblocked_counts_Full$geneIDs<- gsub("mir", "miR", unblocked_counts_Full$geneIDs)
        unblocked_counts_Full_collapsed<-unblocked_counts_Full %>% group_by(geneIDs) %>% summarise_each(funs(sum)) %>% as.data.frame
        rownames(unblocked_counts_Full_collapsed)<-unblocked_counts_Full_collapsed$geneIDs
        unblocked_counts_Full_collapsed<- unblocked_counts_Full_collapsed[,2:ncol(unblocked_counts_Full_collapsed)]
        
#Paired Dataset:
        blocked_counts_paired$geneIDs<- sub("^([^-]*-[^-]*-[^-]*).*", "\\1", rownames(blocked_counts_paired))
        blocked_counts_paired$geneIDs<- gsub("mir", "miR", blocked_counts_paired$geneIDs)
        blocked_counts_paired_collapsed<-blocked_counts_paired %>% group_by(geneIDs) %>% summarise_each(funs(sum)) %>% as.data.frame
        rownames(blocked_counts_paired_collapsed)<-blocked_counts_paired_collapsed$geneIDs
        blocked_counts_paired_collapsed<- blocked_counts_paired_collapsed[,2:ncol(blocked_counts_paired_collapsed)]
        
        unblocked_counts_paired$geneIDs<- sub("^([^-]*-[^-]*-[^-]*).*", "\\1", rownames(unblocked_counts_paired))
        unblocked_counts_paired$geneIDs<- gsub("mir", "miR", unblocked_counts_paired$geneIDs)
        unblocked_counts_paired_collapsed<-unblocked_counts_paired %>% group_by(geneIDs) %>% summarise_each(funs(sum)) %>% as.data.frame
        rownames(unblocked_counts_paired_collapsed)<-unblocked_counts_paired_collapsed$geneIDs
        unblocked_counts_paired_collapsed<- unblocked_counts_paired_collapsed[,2:ncol(unblocked_counts_paired_collapsed)]
```


#2: Plot the number of species passing count thresholds for blocked and unblocked datasets (full cohort)


Get number of species passing thresholds, then plot
```{r}
#Set desired thresholds
thresholds<-seq(0,500,50)
NumberInEachGroup<- length(thresholds)*(89+1) #used for plotting later

#Get counts passing threshold for unblocked samples, then transform data- 4 columns: ID, threshold, #of species that passed that threshold, and blocker status
        unblockedPassThreshold<- as.data.frame(apply(unblocked_counts_Full_collapsed, 2, 
                                                     CountsPassingThreshold, thresholds))
        rownames(unblockedPassThreshold)<- thresholds
        
        #add 4th column: mean of all samples
        unblockedPassThreshold$Mean<- rowMeans(unblockedPassThreshold)
        unblockedPassThreshold<- as.data.frame(t(unblockedPassThreshold))
        unblockedPassThreshold$ID<- rownames(unblockedPassThreshold)
        
        #reformat- melt
        unblockedPassThreshold<- melt(unblockedPassThreshold)
        colnames(unblockedPassThreshold)<- c("ID", "threshold","speciesPassingThreshold")
        unblockedPassThreshold$status<- rep("unblocked", length(thresholds) * (ncol(unblocked_counts_Full_collapsed)+1))
        unblockedPassThreshold$threshold<- as.integer(as.character((unblockedPassThreshold$threshold)))


#Get counts passing threshold for blocked samples, then transform data- 4 columns: ID, threshold, #of species that passed that threshold, and blocker status
        blockedPassThreshold<- as.data.frame(apply(blocked_counts_Full_collapsed, 2, 
                                                     CountsPassingThreshold, thresholds))
        rownames(blockedPassThreshold)<- thresholds
        
        #add 4th column: mean of all samples
        blockedPassThreshold$Mean<- rowMeans(blockedPassThreshold)
        blockedPassThreshold<- as.data.frame(t(blockedPassThreshold))
        blockedPassThreshold$ID<- rownames(blockedPassThreshold)
        
        #reformat- melt
        blockedPassThreshold<- melt(blockedPassThreshold)
        colnames(blockedPassThreshold)<- c("ID", "threshold","speciesPassingThreshold")
        blockedPassThreshold$status<- rep("blocked", length(thresholds) * (ncol(blocked_counts_Full_collapsed)+1))
        blockedPassThreshold$threshold<- as.integer(as.character((blockedPassThreshold$threshold)))

#combine into one
passThreshold<-rbind(unblockedPassThreshold,blockedPassThreshold)
passThreshold$status<-factor(passThreshold$status, levels=c("unblocked", "blocked"))
test<-passThreshold[passThreshold$status=="unblocked",]

#Plot
p<-ggplot(passThreshold, aes(x=threshold, y=speciesPassingThreshold, color=status))+
        geom_point(position="jitter",stat="identity",size=2,alpha=0.2)+
        scale_color_manual(values=c("turquoise3","firebrick3"))+
        theme_classic()+
        scale_x_continuous(breaks=seq(0,500,50))+
        scale_y_continuous(breaks=seq(0,600,100))+
        xlab("Counts threshold")+
        ylab("Number of miRs passing counts threshold")+
        theme(axis.text.x = element_text(size=12,face="bold",color="black"),
              axis.text.y = element_text(size=12,face="bold",color="black"),
              axis.title.x = element_text(size=14,face="bold",color="black"),
              axis.title.y = element_text(size=14,face="bold",color="black"),
              legend.title=element_blank(),
              legend.text = element_text(size=14, face="bold"),
              legend.position=c(.85,.85),
              legend.background = element_rect(fill=NA, size=1, 
                                         linetype="solid", color="black"),
              panel.border = element_rect(colour = "black", fill=NA, size=2))
        


means<- passThreshold[passThreshold$ID=="Mean",]
```


#3: For paired cohort: plot the change in blocked/unblocked number of species passing count thresholds
```{r}
#Set desired thresholds
thresholds<-seq(0,500,100)
NumberInEachGroup<- length(thresholds)*(89+1) #used for plotting later

#Read in paired samples- From BH. 
pairedSamples<-read.csv(paste(wd, "PairedSamples_LargeCohort_FromBH.csv", sep=""), row.names = "UniqueID")
pairedSamples<-pairedSamples[,-1]

#Change column names of counts matrix from sample --> subject. 
unblockedSubject<- as.data.frame(t(unblocked_counts_paired_collapsed))
blockedSubject<- as.data.frame(t(blocked_counts_paired_collapsed))

# Merge with subject info --> make that the rowname --> retranspose
unblockedSubject<-merge(unblockedSubject, pairedSamples, by=0)[,!(colnames(unblockedSubject)=="blocker"),]
blockedSubject<-merge(blockedSubject, pairedSamples, by=0)[,!(colnames(blockedSubject)=="blocker"),]

rownames(unblockedSubject)<-unblockedSubject$phenoId
rownames(blockedSubject)<-blockedSubject$phenoId

unblockedSubject<-unblockedSubject[,-1]
blockedSubject<-blockedSubject[,-1]

unblockedSubject<-as.data.frame(t(unblockedSubject))
blockedSubject<-as.data.frame(t(blockedSubject))

#Convert counts from factor -->integer
unblockedSubject_int<-as.data.frame(apply(unblockedSubject, 2, as.integer))
blockedSubject_int<-as.data.frame(apply(blockedSubject, 2, as.integer))
rownames(unblockedSubject_int)<- rownames(unblockedSubject)
rownames(blockedSubject_int)<- rownames(blockedSubject)
unblockedSubject<-na.omit(unblockedSubject_int)
blockedSubject<-na.omit(blockedSubject_int)

#Get counts passing threshold for unblocked samples, then transform data- 4 columns: ID (A/b/c/d), threshold, #of species that passed that threshold, and blocker status
        unblockedPassThreshold<- as.data.frame(apply(unblockedSubject, 2, 
                                                     CountsPassingThreshold, thresholds))
        rownames(unblockedPassThreshold)<- thresholds
        
        #add 4th column: mean of all samples
        unblockedPassThreshold$Mean<- rowMeans(unblockedPassThreshold)
        unblockedPassThreshold<- as.data.frame(t(unblockedPassThreshold))
        unblockedPassThreshold$ID<- rownames(unblockedPassThreshold)
        
        #reformat- melt
        unblockedPassThreshold<- melt(unblockedPassThreshold)
        colnames(unblockedPassThreshold)<- c("ID", "threshold","speciesPassingThreshold")
        unblockedPassThreshold$status<- rep("unblocked", NumberInEachGroup)
        unblockedPassThreshold$threshold<- as.integer(as.character((unblockedPassThreshold$threshold)))

#Get counts passing threshold for blocked samples, then transform data- 4 columns: ID (A/b/c/d), threshold, #of species that passed that threshold, and blocker status
        blockedPassThreshold<- as.data.frame(apply(blockedSubject, 2, 
                                                     CountsPassingThreshold, thresholds))
        rownames(blockedPassThreshold)<- thresholds
        
        #add 4th column: mean of all samples
        blockedPassThreshold$Mean<- rowMeans(blockedPassThreshold)
        blockedPassThreshold<- as.data.frame(t(blockedPassThreshold))
        blockedPassThreshold$ID<- rownames(blockedPassThreshold)
        
        #reformat- melt
        blockedPassThreshold<- melt(blockedPassThreshold)
        colnames(blockedPassThreshold)<- c("ID", "threshold","speciesPassingThreshold")
        blockedPassThreshold$status<- rep("blocked", NumberInEachGroup)
        blockedPassThreshold$threshold<- as.integer(as.character((blockedPassThreshold$threshold)))

#combine into one
passThreshold<-rbind(unblockedPassThreshold,blockedPassThreshold)

#reformat so that you have 4 columns: ID, threshold, #species passing in noblocker, #species passing in blocker
PT<-data.frame(passThreshold$ID[1:NumberInEachGroup],
               passThreshold$threshold[1:NumberInEachGroup],
               passThreshold$speciesPassingThreshold[1:NumberInEachGroup],
               passThreshold$speciesPassingThreshold[(NumberInEachGroup+1):nrow(passThreshold)])
colnames(PT)<- c("ID", "threshold", "SpeciesPT_noblocker", "SpeciesPT_blocker")

#now need to subtract each- i.e., sample X at threshold 1 with no blocker vs with blocker. get those values for all, then plot. 
df<- as_tibble(PT)
df<-df%>% group_by(ID) %>% mutate(BlockerMinusNoBlocker=SpeciesPT_blocker-SpeciesPT_noblocker) %>% as.data.frame()
df$Mean<- rep(c(rep("Sample",89),"Mean"),length(thresholds))
df$Mean<- as.factor(df$Mean)

Samplecolors<- c("goldenrod1","violetred3", "skyblue3", "darkolivegreen4")

#plot
p<- ggplot(df, aes(x=threshold, y=BlockerMinusNoBlocker, color=ID))+
  geom_line(aes(linetype = Mean, group=ID, size=Mean))+
  scale_linetype_manual(values= c("solid", "dashed"),guide=FALSE)+
  scale_size_manual(values=c("Sample"=1.2,"Mean"=2), guide=FALSE)+
  #geom_point(size=1.3)+
  xlab("Count threshold")+
  ylab("n additional species detected in blocked libraries")+
  #scale_y_continuous(breaks=seq(0, 250, 50), limits=c(0,250))+
  scale_x_continuous(breaks=seq(0, max(df$threshold), 50))+
  #scale_color_manual(values=c("goldenrod1","violetred3", "skyblue3", "darkolivegreen4", "black"))+
  theme_classic()+
  guides(color = guide_legend(override.aes = list(size = 2)))+
  theme(axis.text.y   = element_text(size=12, colour = "black", face= "bold"),
        axis.title.y  = element_text(size=16, colour = "black", face= "bold"),
        axis.text.x   = element_text(size=12, colour = "black", face= "bold"),
        axis.title.x  = element_text(size=16, colour = "black", face= "bold"),
        panel.border = element_rect(colour = "black", fill=NA, size=2),
        legend.title=element_blank(),
        legend.position = c(.9,.85),
        legend.background = element_rect(fill=NA, size=1, linetype="solid", color="black"),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"),
        legend.text=element_text(size=14, face="bold"))
p
```

This plot not used- only actually will use the mean. Will be combined with the pilot graph, same analysis. Export just the mean data here to be used with the pilot:
```{r}
means<-df[df$ID=="Mean",]
write.csv(means, paste(wd, "LargePairedCohort_SpeciesPassingThresholds_MeansOnly.csv",sep=""))
```

```{r}
sessionInfo()
```

