---
title: "Repeating analysis with Mature only"
output: html_notebook
---
libraries
```{r}
suppressPackageStartupMessages(library(metaseqR))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
```

All DESeq2 analyses will be run in "DESeq2Analyses_Pairwise". Other analyses that use raw counts, not a deseq2 object, will be run here- same code as older files (BlockersDecrease, PotentialOfftargets, Improvements), but using MATURE counts as input instead of MATURE + ISOMIRS

Read in data. Select just Mature.

```{r}
wd<- "//Cifs2/rcdata$/Channing_miRNA_Blocker_Test/"
countsInput<- read.delim(paste(wd, "TGC_Data/mirna-blocker_bs_mature-isomirs_raw.txt", sep =""))

#reformat counts data
counts<- countsInput
rownames(counts)<- counts$ID
counts<- counts[,c(4:(ncol(counts)))]

#Select just mature miRNAs. let7 needs to be searched for separately
MatureCounts<- counts[grep("R", rownames(counts)),]
let7<- counts[grep("let", rownames(counts)),]
MatureCounts<- rbind(MatureCounts, let7)
counts<- MatureCounts
```


#Percent of reads mapping to targets in samples without blocker vs with blocker: Collapsing technical replicates

First, collapse technical replicates. Later I'll do this with DESeq2, but here, just averaging counts across technical replicates for A and B samples
```{r}
#Get mean of A/B samples (Collapsing technical replicates)
CollapsedCounts<- counts
CollapsedCounts$A.0_Collapsed<- as.integer(rowMeans(counts[,1:3]))
CollapsedCounts$B.0_Collapsed<- as.integer(rowMeans(counts[,4:6]))
CollapsedCounts$A.1_Collapsed<- as.integer(rowMeans(counts[,9:11]))
CollapsedCounts$B.1_Collapsed<- as.integer(rowMeans(counts[,12:14]))
CollapsedCounts<- CollapsedCounts[,c(17,18,7,8,19,20,15,16)]

colnames(CollapsedCounts)<- c("A.0", "B.0", "C.0", "D.0", "A.1", "B.1", "C.1", "D.1")
```

Next, get number of reads mapping to any of the targets and divide by the total number of reads
```{r}
#Get list of all targets (and very closely related miRNAs)
target_92a<- as.character(rownames(counts[grep("-92a", rownames(counts)),]))
target_92b<- as.character(rownames(counts[grep("-92b", rownames(counts)),]))
target_451a<- as.character(rownames(counts[grep("-451a", rownames(counts)),]))
target_451b<- as.character(rownames(counts[grep("-451b", rownames(counts)),]))
target_486<- as.character(rownames(counts[grep("-486", rownames(counts)),]))
Alltargets<- c(target_92a, target_92b, target_451a, target_451b, target_486)

#get total counts for these targets in all samples
targetCounts<- CollapsedCounts[rownames(CollapsedCounts)%in% Alltargets,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(CollapsedCounts) *100)
PercentMapToTargets$ID<- rownames(PercentMapToTargets)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("NoBlocker", 4), rep("Blocker", 4))
PercentMapToTargets
```


Finally, visualize this change for all 4 samples from blocker to no blocker
```{r}
ggplot(PercentMapToTargets, aes(y=PercentMapToTarget, x=ID, color=BlockerStatus, fill=BlockerStatus))+
         geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  labs(title = "Percent of Total Reads that Map to Targets",
       subtitle = "Technical Replicates collapsed, Target Families Included")+
  scale_y_continuous(name="percentofTotal", breaks=c(seq(0,100, 20)), limit=c(0,100), labels=c(seq(0,100,20)))
 
             
```



```{r}
mean((PercentMapToTargets$PercentMapToTarget[1:4]-PercentMapToTargets$PercentMapToTarget[5:8])/PercentMapToTargets$PercentMapToTarget[1:4])*100
#On average, using the blocker decreases percent mapping to targets by 80%
```


```{r}
#Percent that map to actual targets
ActualTargets<- c('hsa-miR-486-5p','hsa-miR-92a-3p','hsa-miR-451a')

#get total counts for these targets in all samples
targetCounts<- CollapsedCounts[rownames(CollapsedCounts)%in% ActualTargets,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(CollapsedCounts) *100)
PercentMapToTargets$ID<- rownames(PercentMapToTargets)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("NoBlocker", 4), rep("Blocker", 4))
PercentMapToTargets

ggplot(PercentMapToTargets, aes(y=PercentMapToTarget, x=ID, color=BlockerStatus, fill=BlockerStatus))+
         geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  labs(title = "Percent of Total Reads that Map to Targets",
       subtitle = "Technical Replicates collapsed, ONLY actual Targets Included")+
  scale_y_continuous(name="percentofTotal", breaks=c(seq(0,100, 20)), limit=c(0,100), labels=c(seq(0,100,20)))
```

```{r}
mean((PercentMapToTargets$PercentMapToTarget[1:4]-PercentMapToTargets$PercentMapToTarget[5:8])/PercentMapToTargets$PercentMapToTarget[1:4])*100
#On average, using the blocker decreases percent mapping to targets by 80%
```

#Look at the number of reads mapping to species in the same family as the targets but NOT the actual targets themselves:

```{r}
NotTargets<- Alltargets[!(Alltargets %in% ActualTargets)]

#get total counts for these targets in all samples
targetCounts<- CollapsedCounts[rownames(CollapsedCounts)%in% NotTargets,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(CollapsedCounts) *100)
PercentMapToTargets$ID<- rownames(PercentMapToTargets)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("NoBlocker", 4), rep("Blocker", 4))
PercentMapToTargets

ggplot(PercentMapToTargets, aes(y=PercentMapToTarget, x=ID, color=BlockerStatus, fill=BlockerStatus))+
         geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  labs(title = "Percent of Total Reads that Map to Targets",
       subtitle = "Technical Replicates collapsed, ONLY actual Targets Included")+
  scale_y_continuous(name="percentofTotal", breaks=c(seq(0,100, 20)), limit=c(0,100), labels=c(seq(0,100,20)))
```


#Looking at just mature still- except for 451a

mir-451a has a very low expression in its mature form, but very high in its isomir form (see BlockersDecrease). Here, I'm considering the isomir form of 451a + 92a and 486 (exact targets)

```{r}
mir451<- countsInput[countsInput$ID =="hsa-mir-451a",]
rownames(mir451)<- mir451$ID
mir451<- mir451[,4:ncol(mir451)]

counts<- rbind(counts, mir451)

#Get mean of A/B samples (Collapsing technical replicates)
CollapsedCounts<- counts
CollapsedCounts$A.0_Collapsed<- as.integer(rowMeans(counts[,1:3]))
CollapsedCounts$B.0_Collapsed<- as.integer(rowMeans(counts[,4:6]))
CollapsedCounts$A.1_Collapsed<- as.integer(rowMeans(counts[,9:11]))
CollapsedCounts$B.1_Collapsed<- as.integer(rowMeans(counts[,12:14]))
CollapsedCounts<- CollapsedCounts[,c(17,18,7,8,19,20,15,16)]

colnames(CollapsedCounts)<- c("A.0", "B.0", "C.0", "D.0", "A.1", "B.1", "C.1", "D.1")

#Percent that map to actual targets
ActualTargets<- c('hsa-miR-486-5p','hsa-miR-92a-3p','hsa-mir-451a')

#get total counts for these targets in all samples
targetCounts<- CollapsedCounts[rownames(CollapsedCounts)%in% ActualTargets,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(CollapsedCounts) *100)
PercentMapToTargets$ID<- rownames(PercentMapToTargets)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("NoBlocker", 4), rep("Blocker", 4))
PercentMapToTargets

ggplot(PercentMapToTargets, aes(y=PercentMapToTarget, x=ID, color=BlockerStatus, fill=BlockerStatus))+
         geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  labs(title = "Percent of Total Reads that Map to Targets",
       subtitle = "Technical Replicates collapsed, isomir form of 451a + other 2 targets")+
  scale_y_continuous(name="percentofTotal", breaks=c(seq(0,100, 20)), limit=c(0,100), labels=c(seq(0,100,20)))

```

```{r}
mean((PercentMapToTargets$PercentMapToTarget[1:4]-PercentMapToTargets$PercentMapToTarget[5:8])/PercentMapToTargets$PercentMapToTarget[1:4])*100
#On average, using the blocker decreases percent mapping to targets by 87%
```



#Next main section: graphing log2 of cpm

```{r}
#calculating CPM
cpmInput<- MatureCounts

#calculate factor to multiply all counts for to convert to CPM
totalCounts<- colSums(cpmInput)
TotalCountsPM<- 1000000/totalCounts

#Convert all counts to CPM
CPM<-cpmInput*TotalCountsPM
#get rid of any miRNAs that have a 0 counts for any sample- can't take log2 of it
CPMFinal<- CPM[apply(CPM, 1, function(row) all(row !=0 )),]

#log2 of CPM
CPMLog<- log2(CPMFinal)

#Get miRNA info. Need: whether or not it was targeted + whether or not it was DE
#Add in target info- add column for whether or not the species was targeted
targets<- c('hsa-miR-486-5p','hsa-miR-92a-3p','hsa-miR-451a')
CPMLog$target<- rownames(CPMLog) %in% targets

#Add in DE info- using DESeq2 results from targets included analysis. Import data here.
resOrdered<- read.csv(paste(wd, "DESeq2Results/MatureOnly_TargetsIncluded_Pairwise_DESeq2Results_103119.csv", sep =""))
rownames(resOrdered)<- resOrdered$X
resOrdered<- resOrdered[,-1]
sig<- resOrdered[resOrdered$padj<0.05,]
CPMLog$Sig<- rownames(CPMLog) %in% rownames(sig)

#Add in third column, used to color points in graph. Add together target/sig columns. If it's a target and sig, it'll be 2, if it's just sig it'll be 1, if it's neither it'll be 0.
TargetSig<- CPMLog[,17:18]
CPMLog$Color<- as.factor(rowSums(TargetSig))
CPMLog$Color<- gsub(2, "Target", CPMLog$Color)
CPMLog$Color<- gsub(1, "DE", CPMLog$Color)
CPMLog$Color<- gsub(0, "Not DE", CPMLog$Color)
```

```{r}
#Looking at sample A- get average for all 3 Inputs
CPMLogA<- CPMLog[,c(1:3, 9:11,19)]
CPMLogA$A.0<- rowMeans(CPMLogA[,1:3])
CPMLogA$A.1<- rowMeans(CPMLogA[,4:6])
CPMLogA<- CPMLogA[,c(8,9,7)]

pA<- ggplot(CPMLogA, aes(x=A.0, y=A.1, color=Color))+
  geom_point()+
  geom_abline(slope=1, color="red")+
  scale_colour_manual(values=c("red","black", "blue"))+
  theme_classic()
 
pA
```
```{r}
#Looking at sample B- get average for all 3 Inputs
CPMLogB<- CPMLog[,c(4:6, 12:14,19)]
CPMLogB$B.0<- rowMeans(CPMLogB[,1:3])
CPMLogB$B.1<- rowMeans(CPMLogB[,4:6])
CPMLogB<- CPMLogB[,c(8,9,7)]

pB<- ggplot(CPMLogB, aes(x=B.0, y=B.1, color=Color))+
  geom_point()+
  geom_abline(slope=1, color="red")+
  scale_colour_manual(values=c("red","black", "blue"))+
  theme_classic()
 
pB
```


```{r}
#For sample C
CPMLogC<- CPMLog[,c(7, 15,19)]
colnames(CPMLogC)<- c("C.0", "C.1", "Color")

pC<- ggplot(CPMLogC, aes(x=C.0, y=C.1, color=Color))+
  geom_point()+
  geom_abline(slope=1, color="red")+
  scale_colour_manual(values=c("red","black", "blue"))+
  theme_classic()
 
pC
```

```{r}
#For sample D
CPMLogD<- CPMLog[,c(7, 15,19)]
colnames(CPMLogD)<- c("D.0", "D.1", "Color")

pD<- ggplot(CPMLogD, aes(x=D.0, y=D.1, color=Color))+
  geom_point()+
  geom_abline(slope=1, color="red")+
  scale_colour_manual(values=c("red","black", "blue"))+
  theme_classic()
 
pD
```

```{r}
#Averaging values across al 4 samples
CPMLogAll<- CPMLog
CPMLogAll$NoBlocker<- rowMeans(CPMLogAll[,1:8])
CPMLogAll$Blocker<- rowMeans(CPMLogAll[,9:16])
CPMLogAll<- CPMLogAll[,c(20,21,19)]

pMean<- ggplot(CPMLogAll, aes(x=NoBlocker, y=Blocker, color=Color))+
  geom_point()+
  geom_abline(slope=1, color="red")+
  scale_colour_manual(values=c("red","black", "blue"))+
  theme_classic()
 
pMean
```

Putting A-D together on one plot:
```{r}
grid.arrange(pA, pB, pC, pD, nrow=2)
```




#Benefits of blocker: better detection of miRNAs with low counts

Want to look at the number of species that pass a certain numerical threshold for counts- i.e., how many species get at least 100 counts for blocker vs no blocker?

However- big problem: the blocker samples have about 50% of the total reads that no blocker do. So the no blocker will have a much better detection of low level expression- during LC, they would have had a much better chance of picking up lowly expressed miRNAs

So we can't really just normalize- that doesn't quite account for this. 

The way that meyers et al does it (and for much of their paper- might be a good idea to replicate all the other analyses using this type of normalized counts) is to "downsample" their data

This is meant to simulate having the same number of reads/identical LC

They downsample to 6mill reads- I'm downsampling to about 5 million reads, for no reason at all


#Performing downsampling
```{r}
SampledCounts<- downsample.counts(MatureCounts, seed=42)
write.csv(SampledCounts, paste(wd, "OtherResults/DownSampledCounts_MatureOnly_metaseqR_110119.csv", sep =""))

```

#Using the subsampled count matrix generated above, recreate figure 3 from meyers et al: looking at the change in the # of miRNAs that pass a variety of count thresholds for blocked vs unblocked samples. 

Kind of different to what they're plotting, but same idea

```{r}
#read in data
SampledCounts<- read.csv(paste(wd, "OtherResults/DownSampledCounts_MatureOnly_metaseqR_110119.csv", sep =""))
rownames(SampledCounts)<- SampledCounts$X
SampledCounts<- SampledCounts[,-1]

#Collapse technical replicates (by just taking mean of counts)
unblockedCounts<- as.data.frame(SampledCounts[,1:8])
unblockedCounts$ACollapsed<- rowMeans(unblockedCounts[,1:3])
unblockedCounts$BCollapsed<- rowMeans(unblockedCounts[,4:6])
unblockedCounts<-unblockedCounts[,c(9,10,7,8)]

#get mean counts across all 4 samples
unblockedCountsMeans<- as.data.frame((rowMeans(unblockedCounts)))

#number that pass thresholds:
unblockedPassThresholds<- data.frame(threshold=c(1,5,10,50,100,500), countsPassingThreshold=c(sum(unblockedCountsMeans>1),sum(unblockedCountsMeans>5),sum(unblockedCountsMeans>10),sum(unblockedCountsMeans>50),sum(unblockedCountsMeans>100),sum(unblockedCountsMeans>500)))

#blocked: collapse technical replicates (by just taking mean of counts)
blockedCounts<- as.data.frame(SampledCounts[,9:16])
blockedCounts$ACollapsed<- rowMeans(blockedCounts[,1:3])
blockedCounts$BCollapsed<- rowMeans(blockedCounts[,4:6])
blockedCounts<-blockedCounts[,c(9,10,7,8)]

#get mean counts across all 4 samples
blockedCountsMeans<- as.data.frame((rowMeans(blockedCounts)))

#number that pass thresholds:
blockedPassThresholds<- data.frame(threshold=c(1,5,10,50,100,500), countsPassingThreshold=c(sum(blockedCountsMeans>1),sum(blockedCountsMeans>5),sum(blockedCountsMeans>10),sum(blockedCountsMeans>50),sum(blockedCountsMeans>100),sum(blockedCountsMeans>500)))

#combine into one, adding in an extra column for blocker status
unblockedPassThresholds$status<- rep("unblocked",6)
blockedPassThresholds$status<- rep("blocked",6)
passThreshold<-rbind(unblockedPassThresholds,blockedPassThresholds)

#plot
p<- ggplot(passThreshold, aes(x=threshold, y=countsPassingThreshold, color=status))+
             geom_line(size=2)+
  geom_point(size=4)+
  scale_color_manual(values=c("black", "red"))+
  ggtitle("Number of miRNAs passing a count threshold")+
  theme_classic()
p
```




#doing the above, but not taking the mean. still collapse technical, but keep a/b/c/d info. Add in a 4th column in the final table with sample type. Then subtract the #passing threshold for a/b/c/d individually from blocker vs no blocker, then graph. 

whoowhweeee


```{r}
#write (awful) function that takes a column of counts and returns the number of species that pass the 6 threshold already written into the function, bc i am stupid
CountsPassingThreshold<- function(counts){
  return(c(sum(counts>1),sum(counts>5),sum(counts>10),sum(counts>50),sum(counts>100),sum(counts>500),sum(counts>1000)))
}

#Get counts passing threshold for unblocked samples, then transform data- 4 columns: ID (A/b/c/d), threshold, #of species that passed that threshold, and blocker status
unblockedPassThreshold<- as.data.frame(apply(unblockedCounts, 2, CountsPassingThreshold))
rownames(unblockedPassThreshold)<- c(1,5,10,50,100,500,1000)
colnames(unblockedPassThreshold)<- c("A", "B", "C", "D")
#add 4th column: mean of all samples
unblockedPassThreshold$Mean<- rowMeans(unblockedPassThreshold)
unblockedPassThreshold<- as.data.frame(t(unblockedPassThreshold))
unblockedPassThreshold$ID<- rownames(unblockedPassThreshold)
#reformat- melt
unblockedPassThreshold<- melt(unblockedPassThreshold)
colnames(unblockedPassThreshold)<- c("ID", "threshold","speciesPassingThreshold")
unblockedPassThreshold$status<- rep("unblocked", 35)
unblockedPassThreshold$threshold<- as.integer(as.character((unblockedPassThreshold$threshold)))

#Get counts passing threshold for blocked samples, then transform data- 4 columns: ID (A/b/c/d), threshold, #of species that passed that threshold, and blocker status
blockedPassThreshold<- as.data.frame(apply(blockedCounts, 2, CountsPassingThreshold))
rownames(blockedPassThreshold)<- c(1,5,10,50,100,500,1000)
colnames(blockedPassThreshold)<- c("A", "B", "C", "D")
#add 4th column: mean of all samples
blockedPassThreshold$Mean<- rowMeans(blockedPassThreshold)
blockedPassThreshold<- as.data.frame(t(blockedPassThreshold))
blockedPassThreshold$ID<- rownames(blockedPassThreshold)
#reformat- melt
blockedPassThreshold<- melt(blockedPassThreshold)
colnames(blockedPassThreshold)<- c("ID", "threshold","speciesPassingThreshold")
blockedPassThreshold$status<- rep("blocked", 35)
blockedPassThreshold$threshold<- as.integer(as.character((blockedPassThreshold$threshold)))

#combine into one
passThreshold<-rbind(unblockedPassThreshold,blockedPassThreshold)

#reformat so that you have 4 columns: ID, threshold, #species passing in noblocker, #species passing in blocker
PT<-data.frame(passThreshold$ID[1:35], passThreshold$threshold[1:35],passThreshold$speciesPassingThreshold[1:35], passThreshold$speciesPassingThreshold[36:70])
colnames(PT)<- c("ID", "threshold", "SpeciesPT_noblocker", "SpeciesPT_blocker")

#now need to subtract each- i.e., sample A at threshold 1 with no blocker vs with blocker. get those values for all, then plot. 
df<- as_tibble(PT)
df<-df%>% group_by(ID) %>% mutate(BlockerMinusNoBlocker=SpeciesPT_blocker-SpeciesPT_noblocker)
df<- as.data.frame(df)

#plot
p<- ggplot(df, aes(x=threshold, y=BlockerMinusNoBlocker, color=ID))+
             geom_line(size=1.5)+
  geom_point(size=3)+
  scale_color_manual(values=c("turquoise2", "deeppink1", "goldenrod1", "darkviolet","black"))+
  ggtitle("Number of miRNAs passing a count threshold")+
  theme_classic()
p

```


#Next main section: count variation in 3 A0 vs 3 A1- does blocker lower variation?

```{r}
A0<- MatureCounts[,1:3]
A0Mean<-mean(sd(A0$A.190.0), sd(A0$A.570.0), sd(A0$A.760.0))
A1<- MatureCounts[,9:11]
A1Mean<-mean(sd(A1$A.380.1), sd(A1$A.570.1), sd(A1$A.760.1))


(A0Mean-A1Mean)/A0Mean *100
#65% higher variance in no blocker A than in blocker A
```
So seems to suggest that variance is lowered when you use blocker

Repating with sample B, C, D
```{r}
B0<- MatureCounts[,4:6]
B0Mean<-mean(sd(B0$B.190.0), sd(B0$B.570.0), sd(B0$B.760.0))

B1<- MatureCounts[,12:14]
B1Mean<-mean(sd(B1$B.380.1), sd(B1$B.570.1), sd(B1$B.760.1))

(B0Mean-B1Mean)/B0Mean *100
#86% higher variance in no blocker than in blocker

C0<- MatureCounts[,7]
C0Mean<-sd(C0)

C1<- MatureCounts[,15]
C1Mean<-sd(C1)

(C0Mean-C1Mean)/C0Mean *100
#89% higher variance in no blocker than in blocker

D0<- MatureCounts[,8]
D0Mean<-sd(D0)

D1<- MatureCounts[,16]
D1Mean<-sd(D1)

(D0Mean-D1Mean)/D0Mean *100
#85% higher variance in no blocker than in blocker
```

Plotting these changes in variances
```{r}
df<- data.frame(sample = c("A", "A", "B", "B", "C", "C", "D", "D"), blockerStatus=rep(c("noblocker", "blocker"),4), Variation=c(A0Mean, A1Mean, B0Mean, B1Mean, C0Mean, C1Mean, D0Mean,D1Mean))

p1<-ggplot(df, aes(x=sample, y=Variation, color=blockerStatus, fill=blockerStatus))+
  geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  scale_y_continuous(name="Variation: SD", breaks=c(seq(0,160000,10000)), limit=c(0,160000), labels=c(seq(0,160000,10000)))+
  labs(title="Mean Standard Deviation of Raw Counts in Samples A and B")
p1
```


Repeating the above using the downsampled counts data frame instead
```{r}
A0<- SampledCounts[,1:3]
A0Mean<-mean(sd(A0$A.190.0), sd(A0$A.570.0), sd(A0$A.760.0))
A1<- SampledCounts[,9:11]
A1Mean<-mean(sd(A1$A.380.1), sd(A1$A.570.1), sd(A1$A.760.1))


(A0Mean-A1Mean)/A0Mean *100
#51% higher variance in no blocker A than in blocker A
```
So seems to suggest that variance is lowered when you use blocker

Repating with sample B, C, D
```{r}
B0<- SampledCounts[,4:6]
B0Mean<-mean(sd(B0$B.190.0), sd(B0$B.570.0), sd(B0$B.760.0))

B1<- SampledCounts[,12:14]
B1Mean<-mean(sd(B1$B.380.1), sd(B1$B.570.1), sd(B1$B.760.1))

(B0Mean-B1Mean)/B0Mean *100
#68% higher variance in no blocker than in blocker

C0<- SampledCounts[,7]
C0Mean<-sd(C0)

C1<- SampledCounts[,15]
C1Mean<-sd(C1)

(C0Mean-C1Mean)/C0Mean *100
#57% higher variance in no blocker than in blocker

D0<- SampledCounts[,8]
D0Mean<-sd(D0)

D1<- SampledCounts[,16]
D1Mean<-sd(D1)

(D0Mean-D1Mean)/D0Mean *100
#57% higher variance in no blocker than in blocker
```
```{r}
df<- data.frame(sample = c("A", "A", "B", "B", "C", "C", "D", "D"), blockerStatus=rep(c("noblocker", "blocker"),4), Variation=c(A0Mean, A1Mean, B0Mean, B1Mean, C0Mean, C1Mean, D0Mean,D1Mean))

P2<-ggplot(df, aes(x=sample, y=Variation, color=blockerStatus, fill=blockerStatus))+
  geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  scale_y_continuous(name="Variation: SD", breaks=c(seq(0,160000,10000)), limit=c(0,160000), labels=c(seq(0,160000,10000)))+
  labs(title="Mean Standard Deviation of Downsampled Mature Counts")
P2
```

```{r}
grid.arrange(p1,p2, nrow=1)
```