---
title: "Repeating DESeq/etc using SUBSAMPLED COUNTS rather than Raw Counts- Channing Blocker Test"
output: html_notebook
---

Import libraries
```{r}
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(pheatmap))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(gridExtra))
```


Read in data- subsampled counts
```{r}
wd<- "//Cifs2/rcdata$/Channing_miRNA_Blocker_Test/"
counts<- read.csv(paste(wd, "OtherResults/DownSampledCounts_metaseqR_102219.csv", sep=""))
rownames(counts)<- counts$X
counts<- counts[,-1]
```


#PCA


First, need to run DESeq2 to get the normalized counts for PCA. Not actually looking at DE results
```{r}
#rename rows, removing input info
colnames(counts)<- c(rep("A.0", 3), rep("B.0", 3), "C.0", "D.0", rep("A.1", 3), rep("B.1", 3), "C.1", "D.1")

#create metadata
sampleGroup<- c(rep("NoBlocker", 8), rep("Blocker", 8))
sample<- colnames(counts)
meta<- data.frame(sample, sampleGroup)

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

#Collapse replicates- using internal DESeq2 function
ddsColl<- collapseReplicates(dds, dds$sample)

#Run DESeq2
dds<- DESeq(ddsColl)

#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(dds, blind=FALSE)
plotPCA(vsd, intgroup="sampleGroup")
```

#Percent of reads mapping to targets in samples without blocker vs with blocker: Collapsing technical replicates

First, collapse technical replicates. Later I'll do this with DESeq2, but here, just averaging counts across technical replicates for A and B samples
```{r}
#Get mean of A/B samples (Collapsing technical replicates)
CollapsedCounts<- counts
CollapsedCounts$A.0_Collapsed<- as.integer(rowMeans(counts[,1:3]))
CollapsedCounts$B.0_Collapsed<- as.integer(rowMeans(counts[,4:6]))
CollapsedCounts$A.1_Collapsed<- as.integer(rowMeans(counts[,9:11]))
CollapsedCounts$B.1_Collapsed<- as.integer(rowMeans(counts[,12:14]))
CollapsedCounts<- CollapsedCounts[,c(17,18,7,8,19,20,15,16)]

colnames(CollapsedCounts)<- c("A.0", "B.0", "C.0", "D.0", "A.1", "B.1", "C.1", "D.1")
```

Next, get number of reads mapping to any of the targets and divide by the total number of reads
```{r}
#Get list of all targets (and very closely related miRNAs)
target_92a<- as.character(countsInput[grep("-92a", countsInput$ID),1])
target_92b<- as.character(countsInput[grep("-92b", countsInput$ID),1])
target_451a<- as.character(countsInput[grep("-451a", countsInput$ID),1])
target_451b<- as.character(countsInput[grep("-451b", countsInput$ID),1])
target_486<- as.character(countsInput[grep("-486", countsInput$ID),1])
Alltargets<- c(target_92a, target_92b, target_451a, target_451b, target_486)

#get total counts for these targets in all samples
targetCounts<- CollapsedCounts[rownames(CollapsedCounts)%in% Alltargets,]
SumTargetCounts<- colSums(targetCounts)

#Get percentage of total counts that map to targets
PercentMapToTargets<-as.data.frame(SumTargetCounts/colSums(CollapsedCounts) *100)
PercentMapToTargets$ID<- rownames(PercentMapToTargets)
colnames(PercentMapToTargets)<-c("PercentMapToTarget", "ID")
PercentMapToTargets$BlockerStatus<- c(rep("NoBlocker", 4), rep("Blocker", 4))
PercentMapToTargets
```
These values are very similar to using raw counts, not downsampled- only differ in decimal place. This is exactly what we would expect- downsampling shouldn't change the relative proportions of species for HIGHLY EXPRESSED species


Finally, visualize this change for all 4 samples from blocker to no blocker
```{r}
ggplot(PercentMapToTargets, aes(y=PercentMapToTarget, x=ID, color=BlockerStatus, fill=BlockerStatus))+
         geom_bar(stat="identity", position="dodge")
             
```

Skipping the rest of these plots, as this one really doesn't differ at all from raw counts.


#Back to DESeq2: Analyzing DESeq2 results from dds object created previously

```{r}
#Looking at results
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
#14 DE miRNAs- one fewer than using raw counts!

rownames(resOrdered[resOrdered$padj<0.05,])

#write.csv(resOrdered, paste(wd, "DESeq2Results/DESeq2Results_TargetsIncluded_ChanningBlockerPilot.csv", sep = ""))
```

hsa-mir-6837 is missing here and is DE when using raw counts. Looking at counts for this species-

```{r}
#subsampled counts
counts[rownames(counts)=="hsa-mir-6837",]
```


```{r}
#raw counts
countsInput<- read.delim(paste(wd, "TGC_Data/mirna-blocker_bs_mature-isomirs_raw.txt", sep =""))
rownames(countsInput)<- countsInput$ID
countsInput<- countsInput[,4:ncol(countsInput)]
countsInput[rownames(countsInput)=="hsa-mir-6837",]
```

So subsampling decreased the counts for this species very slightly in the blocker sample, enough to make it not DE. mean count for blocked samples of 12.4 vs 16.8

How close was it to being DE?
```{r}
resOrdered[rownames(resOrdered)=="hsa-mir-6837",]
```

Pretty far- p=0.5

Preliminary analysis of potential off target effects: looking at padj <0.1
```{r}
sig<- resOrdered[resOrdered$padj<0.1,]
rownames(sig)
```

################################
#Potential Off target Effects
################################

#Running DESeq2: remove all targets (and their closely related miRNAs), collapse replicates, and run DESeq2. 

If there are minimal off target effects, we shouldn't see hardly any DE miRNAs

First, look at PCA analysis: creating dds object, but not looking at results yet

```{r}
#Get list of all targets
target_92a<- rownames(counts[grep("-92a", rownames(counts)),])
target_92b<- rownames(counts[grep("-92b", rownames(counts)),])
target_451a<- rownames(counts[grep("-451a", rownames(counts)),])
target_451b<- rownames(counts[grep("-451b", rownames(counts)),])
target_486<- rownames(counts[grep("-486", rownames(counts)),])
targets<- c(target_92a, target_92b, target_451a, target_451b, target_486)

#subset countsInput by non-target miRNAs
counts<- counts[!(rownames(counts) %in% targets),]

#rename rows, removing input info
colnames(counts)<- c(rep("A.0", 3), rep("B.0", 3), "C.0", "D.0", rep("A.1", 3), rep("B.1", 3), "C.1", "D.1")

#create metadata
sampleGroup<- c(rep("NoBlocker", 8), rep("Blocker", 8))
sample<- colnames(counts)
meta<- data.frame(sample, sampleGroup)

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

ddsColl<- collapseReplicates(dds, dds$sample)

dds<- DESeq(ddsColl)

#PCA plot for DE with targets excluded
vsd <- varianceStabilizingTransformation(dds, blind=FALSE)
PCAResults<- plotPCA(vsd, intgroup="sampleGroup", returnData=TRUE)
plotPCA(vsd, intgroup="sampleGroup")
```
Similar clustering to using raw counts- very slightly tighter for some samples


#DESeq2 results
```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
#2 DE miRNAs

rownames(resOrdered[resOrdered$padj<0.05,])
#"hsa-mir-6857"  "hsa-miR-25-3p"
```
Same 2 DE as when you use raw counts

Visualizing DE with heatmap:
```{r}
#rld<- rlog(dds)
topgenes<- rownames(resOrdered[1:20,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```



```{r}
counts[rownames(counts)%in% topgenes,]
```





#Recreating graph in Meyers et al- Log of CPM of each miRNA for blocker vs no blocker

Convert counts to cpm --> take log2 --> compare values in blocker vs no blocker

Color points based on whether or not it was a target + whether or not it was sig DE (targets included DESeq2 analysis)

First, looking by individual sample:
```{r}
#calculating CPM
cpmInput<- counts

#calculate factor to multiply all counts for to convert to CPM
totalCounts<- colSums(cpmInput)
TotalCountsPM<- 1000000/totalCounts

#Convert all counts to CPM
CPMFinal<-cpmInput*TotalCountsPM
#get rid of any miRNAs that have a 0 counts for any sample- can't take log2 of it
CPMFinal<- CPMFinal[apply(CPMFinal, 1, function(row) all(row !=0 )),]

#log2 of CPM
CPMLog<- log2(CPMFinal)

#Get miRNA info. Need: whether or not it was targeted + whether or not it was DE
#Add in target info- add column for whether or not the species was targeted
targets<- c('hsa-miR-486-5p','hsa-miR-92a-3p','hsa-mir-451a')
CPMLog$target<- rownames(CPMLog) %in% targets

#Add in DE info- using DESeq2 results from targets included analysis. Import data here.
resOrdered<- read.csv(paste(wd, "DESeq2Results/DESeq2Results_TargetsIncluded_ChanningBlockerPilot.csv", sep =""))
rownames(resOrdered)<- resOrdered$X
resOrdered<- resOrdered[,-1]
sig<- resOrdered[resOrdered$padj<0.05,]
CPMLog$Sig<- rownames(CPMLog) %in% rownames(sig)

#Add in third column, used to color points in graph. Add together target/sig columns. If it's a target and sig, it'll be 2, if it's just sig it'll be 1, if it's neither it'll be 0.
TargetSig<- CPMLog[,17:18]
CPMLog$Color<- as.factor(rowSums(TargetSig))
CPMLog$Color<- gsub(2, "Target", CPMLog$Color)
CPMLog$Color<- gsub(1, "DE", CPMLog$Color)
CPMLog$Color<- gsub(0, "Not DE", CPMLog$Color)
```


```{r}
#Looking at sample A- get average for all 3 Inputs
CPMLogA<- CPMLog[,c(1:3, 9:11,19)]
CPMLogA$A.0<- rowMeans(CPMLogA[,1:3])
CPMLogA$A.1<- rowMeans(CPMLogA[,4:6])
CPMLogA<- CPMLogA[,c(8,9,7)]

pA<- ggplot(CPMLogA, aes(x=A.0, y=A.1, color=Color))+
  geom_point()+
  geom_abline(slope=1, color="red")+
  scale_colour_manual(values=c("red","black", "blue"))+
  theme_classic()
 
pA
```

Similar to raw counts- except that A.1 counts are shifted even farther off the line

```{r}
#Looking at sample B- get average for all 3 Inputs
CPMLogB<- CPMLog[,c(4:6, 12:14,19)]
CPMLogB$B.0<- rowMeans(CPMLogB[,1:3])
CPMLogB$B.1<- rowMeans(CPMLogB[,4:6])
CPMLogB<- CPMLogB[,c(8,9,7)]

pB<- ggplot(CPMLogB, aes(x=B.0, y=B.1, color=Color))+
  geom_point()+
  geom_abline(slope=1, color="red")+
  scale_colour_manual(values=c("red","black", "blue"))+
  theme_classic()
 
pB
```

```{r}
#For sample C
CPMLogC<- CPMLog[,c(7, 15,19)]
colnames(CPMLogC)<- c("C.0", "C.1", "Color")

pC<- ggplot(CPMLogC, aes(x=C.0, y=C.1, color=Color))+
  geom_point()+
  geom_abline(slope=1, color="red")+
  scale_colour_manual(values=c("red","black", "blue"))+
  theme_classic()
 
pC
```

```{r}
#For sample D
CPMLogD<- CPMLog[,c(7, 15,19)]
colnames(CPMLogD)<- c("D.0", "D.1", "Color")

pD<- ggplot(CPMLogD, aes(x=D.0, y=D.1, color=Color))+
  geom_point()+
  geom_abline(slope=1, color="red")+
  scale_colour_manual(values=c("red","black", "blue"))+
  theme_classic()
 
pD
```


```{r}
#Averaging values across al 4 samples
CPMLogAll<- CPMLog
CPMLogAll$NoBlocker<- rowMeans(CPMLogAll[,1:8])
CPMLogAll$Blocker<- rowMeans(CPMLogAll[,9:16])
CPMLogAll<- CPMLogAll[,c(20,21,19)]

pMean<- ggplot(CPMLogAll, aes(x=NoBlocker, y=Blocker, color=Color))+
  geom_point()+
  geom_abline(slope=1, color="red")+
  scale_colour_manual(values=c("red","black", "blue"))+
  theme_classic()
 
pMean
```

Putting A-D together on one plot:
```{r}
grid.arrange(pA, pB, pC, pD, nrow=2)
```


##################################################################
#Benefits of blocker
##################################################################

Only running DESeq2 section for A0 vs B0, not other sections since they were already performed with downsampled count matrix

DESeq2 using A0 vs B0 - here, technical replicates are NOT collapsed
```{r}
#rename rows, removing input info
A0_B0<- counts[,c(1:6)]
colnames(A0_B0)<- c("A.0.1", "A.0.2", "A.0.3", "B.0.1", "B.0.2", "B.0.3")

#create metadata
sampleGroup<- c(rep("NoBlocker", 3), rep("Blocker", 3))
sample<- colnames(A0_B0)
meta<- data.frame(sample, sampleGroup)

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=A0_B0, 
                             colData=meta, 
                             design=~sampleGroup)


dds<- DESeq(dds)

res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
#101 DE miRNAs
noblocker_AvB<- resOrdered
```

DESeq2 using A1 vs B1- here, technical replicates are NOT collapsed
```{r}
#rename rows, removing input info
A1_B1<- counts[,c(9:14)]
colnames(A0_B0)<- c("A.1.1", "A.1.2", "A.1.3", "B.1.1", "B.1.2", "B.1.3")

#create metadata
sampleGroup<- c(rep("NoBlocker", 3), rep("Blocker", 3))
sample<- colnames(A1_B1)
meta<- data.frame(sample, sampleGroup)

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=A1_B1, 
                             colData=meta, 
                             design=~sampleGroup)


dds<- DESeq(dds)

res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
#169 DE miRNAs
blocker_AvB<- resOrdered
```

```{r}
#So that's a....
(169-101)/101
#67% increase in the number of DE genes!!!!!!
```
When using downsampled counts, you get an even greater number of DE miRNAs. Makes sense, right? The blocker gives you more species with really low abundance in the sample, but that improvement is partially masked bc there's more reads for the no blocker samples. When you get rid of that advantage, you can see the blocker improvement even more clearly.

Something that might be interesting to look at- where do these 2 sets of DE miRNAs differ? If the reasoning above is correct, most should be the same and a lot of the ones that differ should have pretty low overall counts
```{r}
#Get sig- considering sig to be .1

noblocker_sig_AvB<-noblocker_AvB[noblocker_AvB$padj<0.1,]
blocker_sig_AvB<-blocker_AvB[blocker_AvB$padj<0.1,]
#145 DE vs 225 DE- 56% improvement

DEInBoth<- rownames(blocker_sig_AvB) %in% rownames(noblocker_sig_AvB) 
#107 in both- that is, 74% of the miRNAs DE in no blocker are also DE in blocker. Seems a little low? Tring with sig set as 0.05 to maybe get rid of some background

noblocker_sig_AvB<-noblocker_AvB[noblocker_AvB$padj<0.05,]
blocker_sig_AvB<-blocker_AvB[blocker_AvB$padj<0.05,]

DEInBoth<- rownames(blocker_sig_AvB) %in% rownames(noblocker_sig_AvB)
#Still about the same- 71%. So you really have some that are DE in no blocker but not in blocker. I guess there's just some inherent variability that you can't get rid of? 


#get miRNAs DE in blocker but not in noblocker
OnlyBlockerDE<- blocker_sig_AvB[!DEInBoth,]

#looking at raw counts of these DE miRNAs
counts_onlyblockersDE<- counts[rownames(counts) %in% rownames(OnlyBlockerDE),]
counts_onlyblockersDE<- counts_onlyblockersDE[,c(1:6,9:14)]
#get the mean for A/B w/ and w/o blocker
counts_onlyblockersDE$A.0Mean<- rowMeans(counts_onlyblockersDE[,1:3])
counts_onlyblockersDE$B.0Mean<- rowMeans(counts_onlyblockersDE[,4:6])
counts_onlyblockersDE$A.1Mean<- rowMeans(counts_onlyblockersDE[,7:9])
counts_onlyblockersDE$B.1Mean<- rowMeans(counts_onlyblockersDE[,10:12])
counts_onlyblockersDE<- counts_onlyblockersDE[,13:16]
counts_onlyblockersDE<- as.data.frame(apply(counts_onlyblockersDE, 2, as.integer))
rownames(counts_onlyblockersDE)<- rownames(counts[rownames(counts) %in% rownames(OnlyBlockerDE),])
head(counts_onlyblockersDE)
```
```{r}
tail(counts_onlyblockersDE,25)
```

When looking at the lowest counts of these miRNAs ONLY DE in blocked AvB, you see that in general, the overall counts are lower for unblocked samples. This supports the idea that the blocker gives better expression of lowly expressed miRNAs, explaining why these lowly expressed miRNAs are able to be detected as DE in blocked but not in blocked.

Looking at average counts for these lowly expressed DE miRNAs:
```{r}
colMeans(tail(counts_onlyblockersDE,25))
```
Way higher for blocked! Starts to normalize to about the same level when you look at the bottom 50. Compared to overall counts for these DE miRNAs- would expect similar counts, maybe higher for blocked:

```{r}
colSums(counts_onlyblockersDE)
```
About ~50% lower counts for blocked- even though counts are normalized to the same total number for all samples. Suggests a better spread of counts across these DE miRNAs, like we saw in the figure plotting #counts passing threshold.


Something else to think about- comparing the list of DE miRNAs between blocked/unblocked, look at which pass a counts threshold- like I did previously.
```{r}
#write (awful) function that takes a column of counts and returns the number of species that pass the 6 threshold already written into the function, bc i am stupid
CountsPassingThreshold<- function(counts){
  return(c(sum(counts>1),sum(counts>5),sum(counts>10),sum(counts>50),sum(counts>75),sum(counts>100),sum(counts>200)))
}

#Get counts passing threshold for DE miRNAs in unblocked samples, then transform data- 4 columns: ID (A/b/mean), threshold, #of species that passed that threshold, and blocker status
unblockedCounts<- counts[rownames(counts) %in% rownames(noblocker_sig_AvB),1:6]
unblockedCounts$Mean<- rowMeans(unblockedCounts)
unblockedCounts<- as.data.frame(unblockedCounts[,7])
unblockedPassThreshold<- as.data.frame(apply(unblockedCounts, 2, CountsPassingThreshold))
rownames(unblockedPassThreshold)<- c(1,5,10,50,100,500,1000)
#add 4th column: mean of all samples
unblockedPassThreshold<- as.data.frame(t(unblockedPassThreshold))
#reformat- melt
unblockedPassThreshold<- melt(unblockedPassThreshold)
colnames(unblockedPassThreshold)<- c("threshold","speciesPassingThreshold")
unblockedPassThreshold$status<- rep("unblocked", 7)
unblockedPassThreshold$threshold<- as.integer(as.character((unblockedPassThreshold$threshold)))
#convert number passing threshold to a percentage- divide by the total number of DE miRNAs
unblockedPassThreshold$speciesPassingThreshold<- unblockedPassThreshold$speciesPassingThreshold/nrow(unblockedCounts) *100

#Get counts passing threshold for DE miRNAs in blocked samples, then transform data- 4 columns: ID (A/b/mean), threshold, #of species that passed that threshold, and blocker status
blockedCounts<- counts[rownames(counts) %in% rownames(blocker_sig_AvB),8:14]
blockedCounts$Mean<- rowMeans(blockedCounts)
blockedCounts<- as.data.frame(blockedCounts[,7])
blockedPassThreshold<- as.data.frame(apply(blockedCounts, 2, CountsPassingThreshold))
rownames(blockedPassThreshold)<- c(1,5,10,50,100,500,1000)
#add 4th column: mean of all samples
blockedPassThreshold<- as.data.frame(t(blockedPassThreshold))
#reformat- melt
blockedPassThreshold<- melt(blockedPassThreshold)
colnames(blockedPassThreshold)<- c("threshold","speciesPassingThreshold")
blockedPassThreshold$status<- rep("blocked", 7)
blockedPassThreshold$threshold<- as.integer(as.character((blockedPassThreshold$threshold)))
#convert number passing threshold to a percentage- divide by the total number of DE miRNAs
blockedPassThreshold$speciesPassingThreshold<- blockedPassThreshold$speciesPassingThreshold/nrow(blockedCounts) *100

PT<- rbind(unblockedPassThreshold, blockedPassThreshold)


#plot
p<- ggplot(PT, aes(x=threshold, y=speciesPassingThreshold, color=status))+
             geom_line(size=1.5)+ geom_point(size=3) +
  scale_color_manual(values=c("darksalmon", "dodgerblue4"))+
  ggtitle("Number of miRNAs passing a count threshold")+
  theme_classic()
p
```


Hmm so not a huge difference. Maybe not the best way to visualize this. Will keep thinking it through...

