---
title: "Pairwise DESeq2 analysis"
output: html_notebook
---
Import libraries
```{r}
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(pheatmap))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(gridExtra))
```


Read in data
```{r}
wd<- "//Cifs2/rcdata$/Channing_miRNA_Blocker_Test/"
countsInput<- read.delim(paste(wd, "TGC_Data/mirna-blocker_bs_mature-isomirs_raw.txt", sep =""))

#reformat counts data
counts<- countsInput
rownames(counts)<- counts$ID
counts<- counts[,c(4:(ncol(counts)))]
```

#Using mature and isomir data for DESeq2 input

```{r}
#rename rows, removing input info
colnames(counts)<- c(rep("A.0", 3), rep("B.0", 3), "C.0", "D.0", rep("A.1", 3), rep("B.1", 3), "C.1", "D.1")

#################
#create metadata#
#################

#set blocker status
sampleGroup<- as.factor(c(rep("NoBlocker", 8), rep("Blocker", 8)))
#this column is used to collapse technical replicates
sample<- colnames(counts)
#this column is to control for sample to sample differences
patient<- as.factor(c(rep("A", 3), rep("B", 3), "C", "D", rep("A", 3), rep("B", 3), "C", "D"))
meta<- data.frame(sample, sampleGroup, patient)

#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

#Collapse replicates- using internal DESeq2 function
ddsColl<- collapseReplicates(dds, dds$sample)

#Create model that takes into account sample differences
ddsMF<- ddsColl
design(ddsMF)<- formula(~patient + sampleGroup)

#run DESeq2
dds<- DESeq(ddsMF)

#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(ddsMF, blind=FALSE)
PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ggtitle("PCA: Blocker vs no Blocker, Mature + Isomirs")
  
p
```


#DESeq2 results
```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.01,])
#55 DE miRNAs- padj <0.05
#31 DE miRNAs- padj <0.01

rownames(resOrdered[resOrdered$padj<0.01,])

```


Visualizing DE with heatmap:
```{r}
rld<- rlog(dds)
topgenes<- rownames(resOrdered[1:75,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```


#Repeating the above analysis, with only mature miRNAs as input instead:

```{r}
#Select just mature miRNAs. let7 needs to be searched for separately
MatureCounts<- counts[grep("R", rownames(counts)),]
let7<- counts[grep("let", rownames(counts)),]
MatureCounts<- rbind(MatureCounts, let7)
counts<- MatureCounts

#rename rows, removing input info
colnames(counts)<- c(rep("A.0", 3), rep("B.0", 3), "C.0", "D.0", rep("A.1", 3), rep("B.1", 3), "C.1", "D.1")

#################
#create metadata#
#################

#set blocker status
sampleGroup<- as.factor(c(rep("NoBlocker", 8), rep("Blocker", 8)))
#this column is used to collapse technical replicates
sample<- colnames(counts)
#this column is to control for sample to sample differences
patient<- as.factor(c(rep("A", 3), rep("B", 3), "C", "D", rep("A", 3), rep("B", 3), "C", "D"))
meta<- data.frame(sample, sampleGroup, patient)

#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

#Collapse replicates- using internal DESeq2 function
ddsColl<- collapseReplicates(dds, dds$sample)

#Create model that takes into account sample differences
ddsMF<- ddsColl
design(ddsMF)<- formula(~patient + sampleGroup)

#run DESeq2
dds<- DESeq(ddsMF)

#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(dds, blind=FALSE)
PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ggtitle("PCA: Blocker vs no Blocker, Mature ONLY")
  
p

```

```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.01,])
#18 DE miRNAs- padj <0.05
#13 DE miRNAs- padj <0.01

rownames(resOrdered[resOrdered$padj<0.05,])
```

```{r}
#Save these off target mIRNAs for later
sigDE_mature<- resOrdered[resOrdered$padj<0.01,]
```

```{r}
rld<- rlog(dds)
topgenes<- rownames(resOrdered[1:20,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```

```{r}
#write to file \
write.csv(resOrdered, paste(wd, "DESeq2Results/MatureOnly_Pairwise_DESeq2Results_103119.csv", sep = "" ))
```

For fun- looking at just sample A here. technical replicates not collapsed. Using only mature.

```{r}
SampleA<- counts[,grep("A", colnames(counts))]

#set blocker status
sampleGroup<- as.factor(c(rep("NoBlocker", 3), rep("Blocker", 3)))
#this column is used to collapse technical replicates
sample<- colnames(SampleA)

meta<- data.frame(sample, sampleGroup)

#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later
dds<- DESeqDataSetFromMatrix(countData=SampleA, 
                             colData=meta, 
                             design=~sampleGroup)


#run DESeq2
dds<- DESeq(dds)

#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(dds, blind=FALSE)
PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ggtitle("PCA: Blocker vs no Blocker")
  
p
```

```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.01,])
#9 DE miRNAs- padj <0.05
#6 DE miRNAs- padj <0.01

rownames(resOrdered[resOrdered$padj<0.05,])
```

```{r}
rld<- rlog(dds)
topgenes<- rownames(resOrdered[1:20,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```

For fun- looking at just sample B here. technical replicates not collapsed. Using only mature.

```{r}
SampleB<- counts[,grep("B", colnames(counts))]

#set blocker status
sampleGroup<- as.factor(c(rep("NoBlocker", 3), rep("Blocker", 3)))
#this column is used to collapse technical replicates
sample<- colnames(SampleB)

meta<- data.frame(sample, sampleGroup)

#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later
dds<- DESeqDataSetFromMatrix(countData=SampleB, 
                             colData=meta, 
                             design=~sampleGroup)


#run DESeq2
dds<- DESeq(dds)

#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(dds, blind=FALSE)
PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ggtitle("PCA: Blocker vs no Blocker")
  
p
```

```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
#6 DE miRNAs

rownames(resOrdered[resOrdered$padj<0.05,])
```

```{r}
rld<- rlog(dds)
topgenes<- rownames(resOrdered[1:20,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```


#Looking at off target effects

Graphing raw counts for all sig DE
```{r}
targetfamilies<- c(Alltargets, "hsa-miR-25-3p")
sigCounts<- MatureCounts[rownames(MatureCounts) %in% rownames(sigDE_mature),]
sigCounts$Mean<- rowMeans(sigCounts)
sigCounts$ID<- rownames(sigCounts)
sigCounts<- sigCounts[order(sigCounts$Mean),]
sigCounts$TargetFamilies<-rownames(sigCounts) %in% targetfamilies

ggplot(sigCounts[,c(17:19)], aes(x=ID, y=Mean,color=TargetFamilies, fill=TargetFamilies))+
  geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  labs(title = "Mean Raw counts of sig DE miRNAs",
       subtitle = "")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_y_continuous(name="Mean raw counts")
```

Graphing log2 fold change for all sig DE

```{r}
df<- data.frame(ID=rownames(sigDE_mature), log2=sigDE_mature$log2FoldChange, targetFamilies= rownames(sigDE_mature) %in% targetfamilies)

#get rid of any miRNAs where the mean count is < 50
sigCounts<- sigCounts[rowMeans(sigCounts[,1:16])>50,]

ggplot(df, aes(x=ID, y=log2,color=targetFamilies, fill=targetFamilies))+
  geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  labs(title = "Log2 fold change of sig DE miRNAs",
       subtitle = "")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_y_continuous(name="log2 fold change")

```

```{r}
#sig de with postiive log fold 2 change
poslog2<-rownames(sigDE_mature[sigDE_mature$log2FoldChange>0,])
POTs<- poslog2[!(poslog2 %in% targetfamilies)]

rowMeans(MatureCounts[rownames(MatureCounts) %in% POTs,])
```


#Remove targets --> Run DESeq2

```{r}
#Get list of all targets
ActualTargets<- c('hsa-miR-486-5p','hsa-miR-92a-3p','hsa-miR-451a')

#subset counts (mature only) by non-target miRNAs
counts<- counts[!(rownames(counts) %in% ActualTargets),]

#################
#create metadata#
#################

#set blocker status
sampleGroup<- as.factor(c(rep("NoBlocker", 8), rep("Blocker", 8)))
#this column is used to collapse technical replicates
sample<- colnames(counts)
#this column is to control for sample to sample differences
patient<- as.factor(c(rep("A", 3), rep("B", 3), "C", "D", rep("A", 3), rep("B", 3), "C", "D"))
meta<- data.frame(sample, sampleGroup, patient)

#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

#Collapse replicates- using internal DESeq2 function
ddsColl<- collapseReplicates(dds, dds$sample)

#Create model that takes into account sample differences
ddsMF<- ddsColl
design(ddsMF)<- formula(~patient + sampleGroup)

#run DESeq2
dds<- DESeq(ddsMF)

#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(ddsMF, blind=FALSE)
PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ggtitle("PCA: Blocker vs no Blocker, Mature ONLY")
  
p
```


Samples still cluster by SAMPLE when you remove targets from the analysis- even for pairwise analysis!

```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.01,])
#17 DE miRNAs padj <0.05
#13 DE miRNAs padj <0.01

rownames(resOrdered[resOrdered$padj<0.01,])
```

```{r}
rld<- rlog(dds)
topgenes<- rownames(resOrdered[1:20,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```




```{r}
#Looking at the off targets from this analysis
sigDE_mature<- resOrdered[resOrdered$padj<0.01,]
targetfamilies<- c(Alltargets, "hsa-miR-25-3p")
sigCounts<- MatureCounts[rownames(MatureCounts) %in% rownames(sigDE_mature),]
sigCounts$Mean<- rowMeans(sigCounts)
sigCounts$ID<- rownames(sigCounts)
sigCounts<- sigCounts[order(sigCounts$Mean),]
sigCounts$TargetFamilies<-rownames(sigCounts) %in% targetfamilies

#get rid of any miRNAs where the mean count is < 50
sigCounts<- sigCounts[rowMeans(sigCounts[,1:16])>50,]

ggplot(sigCounts[,c(17:19)], aes(x=ID, y=Mean,color=TargetFamilies, fill=TargetFamilies))+
  geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  labs(title = "Mean Raw counts of sig DE miRNAs",
       subtitle = "")+
  scale_y_continuous(name="Mean raw counts")+
  theme(axis.text.x = element_text(angle = 90))

```


```{r}
df<- data.frame(ID=rownames(sigDE_mature), log2=sigDE_mature$log2FoldChange, targetFamilies= rownames(sigDE_mature) %in% targetfamilies)

ggplot(df, aes(x=ID, y=log2,color=targetFamilies, fill=targetFamilies))+
  geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  labs(title = "Log2 fold change of sig DE miRNAs",
       subtitle = "")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_y_continuous(name="log2 fold change")
```

#DESeq2: on families

Same as previously, just running for pairwise

```{r}
#Import families
families<- read.delim(paste(wd, "TGC_Data/mirna-blocker_bs_mirna-families_raw.txt", sep =""))
rownames(families)<- families$ID
families<- families[,4:ncol(families)]
```

```{r}
#rename rows, removing input info
colnames(families)<- c(rep("A.0", 3), rep("B.0", 3), "C.0", "D.0", rep("A.1", 3), rep("B.1", 3), "C.1", "D.1")

#################
#create metadata#
#################

#set blocker status
sampleGroup<- as.factor(c(rep("NoBlocker", 8), rep("Blocker", 8)))
#this column is used to collapse technical replicates
sample<- colnames(families)
#this column is to control for sample to sample differences
patient<- as.factor(c(rep("A", 3), rep("B", 3), "C", "D", rep("A", 3), rep("B", 3), "C", "D"))
meta<- data.frame(sample, sampleGroup, patient)

#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later
dds<- DESeqDataSetFromMatrix(countData=families, 
                             colData=meta, 
                             design=~sampleGroup)

#Collapse replicates- using internal DESeq2 function
ddsColl<- collapseReplicates(dds, dds$sample)

#Create model that takes into account sample differences
ddsMF<- ddsColl
design(ddsMF)<- formula(~patient + sampleGroup)

#run DESeq2
dds<- DESeq(ddsMF)

#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(ddsMF, blind=FALSE)
PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ggtitle("PCA: Blocker vs no Blocker, Mature + Isomirs")
  
p
```


```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
#20 DE miRNAs

rownames(resOrdered[resOrdered$padj<0.05,])

```


Visualizing DE with heatmap:
```{r}
#rld<- rlog(dds)
topgenes<- rownames(resOrdered[1:25,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```









#Improvements with blockers- DESeq2 of A0 vs B0 and A1 vs B1

#Next main section: Does using blocker get you more DE- A0 vs B0 compared to A1 vs B1


DESeq2 using A0 vs B0 - here, technical replicates are NOT collapsed

ALSO- PAIRWISE ANALYSIS NOT PERFORMED
Don't want to control for sample differences here, since we're comparing between samples
```{r}
#rename rows, removing input info
A0_B0<- MatureCounts[,c(1:6)]
colnames(A0_B0)<- c("A.0.1", "A.0.2", "A.0.3", "B.0.1", "B.0.2", "B.0.3")

#create metadata
sampleGroup<- c(rep("NoBlocker", 3), rep("Blocker", 3))
sample<- colnames(A0_B0)
meta<- data.frame(sample, sampleGroup)

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=A0_B0, 
                             colData=meta, 
                             design=~sampleGroup)


dds<- DESeq(dds)

res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
#49 DE miRNAs
```

```{r}
rld<- rlog(dds)
topgenes<- rownames(resOrdered[resOrdered$padj<0.05,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```


DESeq2 using A1 vs B1- here, technical replicates are NOT collapsed
```{r}
#rename rows, removing input info
A1_B1<- MatureCounts[,c(9:14)]
colnames(A0_B0)<- c("A.1.1", "A.1.2", "A.1.3", "B.1.1", "B.1.2", "B.1.3")

#create metadata
sampleGroup<- c(rep("NoBlocker", 3), rep("Blocker", 3))
sample<- colnames(A1_B1)
meta<- data.frame(sample, sampleGroup)

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=A1_B1, 
                             colData=meta, 
                             design=~sampleGroup)


dds<- DESeq(dds)

res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
#83 DE miRNAs
```

```{r}
rld<- rlog(dds)
topgenes<- rownames(resOrdered[resOrdered$padj<0.05,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```





#Now using downsampled Counts: repeating a few analyses:

1) mature downsampled counts --> keep targets --> DESeq2
  -PCA
  -graph raw counts/fold change of sig DE
2) mature downsampled counts --> removed targets --> DESeq2
  -PCA
  -graph raw counts/fold change of sig DE
3) log of cpm of mature downsampled counts


#Number 1: mature downsampled --> DESeq2

```{r}
#read in downsampled data
SampledCounts<- read.csv(paste(wd, "OtherResults/DownSampledCounts_MatureOnly_metaseqR_110119.csv", sep =""))
rownames(SampledCounts)<- SampledCounts$X
SampledCounts<- SampledCounts[,-1]
```


DEseq2
```{r}
counts<- SampledCounts

#rename rows, removing input info
colnames(counts)<- c(rep("A.0", 3), rep("B.0", 3), "C.0", "D.0", rep("A.1", 3), rep("B.1", 3), "C.1", "D.1")

#################
#create metadata#
#################

#set blocker status
sampleGroup<- as.factor(c(rep("NoBlocker", 8), rep("Blocker", 8)))
#this column is used to collapse technical replicates
sample<- colnames(counts)
#this column is to control for sample to sample differences
patient<- as.factor(c(rep("A", 3), rep("B", 3), "C", "D", rep("A", 3), rep("B", 3), "C", "D"))
meta<- data.frame(sample, sampleGroup, patient)

#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

#Collapse replicates- using internal DESeq2 function
ddsColl<- collapseReplicates(dds, dds$sample)

#Create model that takes into account sample differences
ddsMF<- ddsColl
design(ddsMF)<- formula(~patient + sampleGroup)

#run DESeq2
dds<- DESeq(ddsMF)

#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(dds, blind=FALSE)
PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ggtitle("PCA: Blocker vs no Blocker, Mature ONLY, downsampled")
  
p

```

Even with downsampled counts, still cluster by blocker status/ a/b vs c/d. Makes sense.

#DESeq2 results
```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
#18 DE miRNAs

rownames(resOrdered[resOrdered$padj<0.05,])

```


Visualizing DE with heatmap:
```{r}
rld<- rlog(dds)
topgenes<- rownames(resOrdered[1:20,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```

```{r}
#Looking at the off targets from this analysis
sigDE_mature<- resOrdered[resOrdered$padj<0.05,]
targetfamilies<- c(Alltargets, "hsa-miR-25-3p")
sigCounts<- SampledCounts[rownames(SampledCounts) %in% rownames(sigDE_mature),]
sigCounts$Mean<- rowMeans(sigCounts)
sigCounts$ID<- rownames(sigCounts)
sigCounts<- sigCounts[order(sigCounts$Mean),]
sigCounts$TargetFamilies<-rownames(sigCounts) %in% targetfamilies
sigCounts<- sigCounts[,c(17:19)]

ggplot(sigCounts, aes(x=ID, y=Mean,color=TargetFamilies, fill=TargetFamilies))+
  geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  labs(title = "Mean Raw counts of sig DE miRNAs",
       subtitle = "")+
  scale_y_continuous(name="Mean raw counts")+
  theme(axis.text.x = element_text(angle = 90))

```


```{r}
df<- data.frame(ID=rownames(sigDE_mature), log2=sigDE_mature$log2FoldChange, targetFamilies= rownames(sigDE_mature) %in% targetfamilies)

ggplot(df, aes(x=ID, y=log2,color=targetFamilies, fill=targetFamilies))+
  geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  labs(title = "Log2 fold change of sig DE miRNAs",
       subtitle = "")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_y_continuous(name="log2 fold change")
```

DEseq2
```{r}
counts<- SampledCounts[!(rownames(SampledCounts) %in% ActualTargets),]

#rename rows, removing input info
colnames(counts)<- c(rep("A.0", 3), rep("B.0", 3), "C.0", "D.0", rep("A.1", 3), rep("B.1", 3), "C.1", "D.1")

#################
#create metadata#
#################

#set blocker status
sampleGroup<- as.factor(c(rep("NoBlocker", 8), rep("Blocker", 8)))
#this column is used to collapse technical replicates
sample<- colnames(counts)
#this column is to control for sample to sample differences
patient<- as.factor(c(rep("A", 3), rep("B", 3), "C", "D", rep("A", 3), rep("B", 3), "C", "D"))
meta<- data.frame(sample, sampleGroup, patient)

#create DESeq2 object- just based on samplegroup (blockerstatus), will add samples to model later
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

#Collapse replicates- using internal DESeq2 function
ddsColl<- collapseReplicates(dds, dds$sample)

#Create model that takes into account sample differences
ddsMF<- ddsColl
design(ddsMF)<- formula(~patient + sampleGroup)

#run DESeq2
dds<- DESeq(ddsMF)

#Normalized (variance stabilized transformation), then run PCA
vsd <- varianceStabilizingTransformation(dds, blind=FALSE)
PCA<-plotPCA(vsd, intgroup="sampleGroup",returnData=TRUE)

p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=4)+
  scale_color_manual(values=c("orangered1","springgreen2"))+
  theme_classic()+
  ggtitle("PCA: Blocker vs no Blocker, Mature ONLY, downsampled, targets removed")
  
p

```

Even with downsampled counts, still cluster by blocker status/ a/b vs c/d. Makes sense.

#DESeq2 results
```{r}
res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
#16 DE miRNAs

rownames(resOrdered[resOrdered$padj<0.05,])

```


Visualizing DE with heatmap:
```{r}
rld<- rlog(dds)
topgenes<- rownames(resOrdered[1:20,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- 0
subtractFromMax<- 1
lengthOut<- 75
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```

```{r}
#Looking at the off targets from this analysis
sigDE_mature<- resOrdered[resOrdered$padj<0.05,]
targetfamilies<- c(Alltargets, "hsa-miR-25-3p")
sigCounts<- SampledCounts[rownames(SampledCounts) %in% rownames(sigDE_mature),]
sigCounts$Mean<- rowMeans(sigCounts)
sigCounts$ID<- rownames(sigCounts)
sigCounts<- sigCounts[order(sigCounts$Mean),]
sigCounts$TargetFamilies<-rownames(sigCounts) %in% targetfamilies
sigCounts<- sigCounts[,c(17:19)]

ggplot(sigCounts, aes(x=ID, y=Mean,color=TargetFamilies, fill=TargetFamilies))+
  geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  labs(title = "Mean Raw counts of sig DE miRNAs",
       subtitle = "")+
  scale_y_continuous(name="Mean raw counts")+
  theme(axis.text.x = element_text(angle = 90))

```

```{r}
df<- data.frame(ID=rownames(sigDE_mature), log2=sigDE_mature$log2FoldChange, targetFamilies= rownames(sigDE_mature) %in% targetfamilies)

ggplot(df, aes(x=ID, y=log2,color=targetFamilies, fill=targetFamilies))+
  geom_bar(stat="identity", position="dodge")+
  theme_classic()+
  labs(title = "Log2 fold change of sig DE miRNAs",
       subtitle = "")+
  theme(axis.text.x = element_text(angle = 90))+
  scale_y_continuous(name="log2 fold change")
```




#benefits: improved DESeq2. Repeating analysis with downsampled

DESeq2 using A0 vs B0 - here, technical replicates are NOT collapsed

ALSO- PAIRWISE ANALYSIS NOT PERFORMED
Don't want to control for sample differences here, since we're comparing between samples
```{r}
#rename rows, removing input info
A0_B0<- SampledCounts[,c(1:6)]
colnames(A0_B0)<- c("A.0.1", "A.0.2", "A.0.3", "B.0.1", "B.0.2", "B.0.3")

#create metadata
sampleGroup<- c(rep("NoBlocker", 3), rep("Blocker", 3))
sample<- colnames(A0_B0)
meta<- data.frame(sample, sampleGroup)

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=A0_B0, 
                             colData=meta, 
                             design=~sampleGroup)


dds<- DESeq(dds)

res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
#41 DE miRNAs
```

```{r}
#rename rows, removing input info
A1_B1<- SampledCounts[,c(9:14)]
colnames(A0_B0)<- c("A.1.1", "A.1.2", "A.1.3", "B.1.1", "B.1.2", "B.1.3")

#create metadata
sampleGroup<- c(rep("NoBlocker", 3), rep("Blocker", 3))
sample<- colnames(A1_B1)
meta<- data.frame(sample, sampleGroup)

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=A1_B1, 
                             colData=meta, 
                             design=~sampleGroup)


dds<- DESeq(dds)

res<- results(dds)
res<- na.omit(res)
resOrdered<- res[order(res$padj),]
nrow(resOrdered[resOrdered$padj<0.05,])
#73 DE miRNAs
```
